---
title: "R Notebook"
output: html_notebook
---
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(here)
library(deSolve)
library(crone)
library(optimParallel)
library(doParallel)
library(doRNG)
library(arrow)
library(stringr)
library(parallel)
library(ggpubr)
```

Notebook for plotting all of the figures for PloS Biology manuscript submission

Guidelines: taken from https://journals.plos.org/plosbiology/s/figures#loc-figure-file-requirements
1. format: tiff
2. max file size: 10 MB
3. text size: Arial, Times, or Symbol font only in 8-12 point
2. figure size: Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).

# overall manuscript setup
1. Fig 1: model schematic
2. Fig 2. best single and co-infection infection cue [monocue] -------> supplmentary table of parameters and fitness
3. Fig 3. dual cue optimization [monocue] -------> supplementary table of flexible optimization
4. Fig 4. heterocue competition [heterocue] 
5. Fig 5. partitioning single cue success and consequences of bottom+top down regulation [monocue]
6. Fig 6. consequences of adopting best cue [monocue]

#-----------------------------------#
best single and co-infection cue
#-----------------------------------#
Figure displaying the reaction norms of best single and co-infection.

#------- optimal cue reaction norm -----------#
# read data
```{r}
# single infection dynamics, reaction norms, and rugs
si_dyn.df <- read_parquet(here("data/si_dyn/si_dyn_30.parquet")) 
si_rn.df <- read_parquet(here("data/si_dyn/si_rn.parquet"))
si_rug.df <- read_parquet(here("data/si_dyn/si_rug.parquet"))

# co-infection dynamics, reaction norms, and rugs
ci_dyn.df <- read_parquet(here("data/ci_dyn/ci_dyn.parquet"))
ci_rn.df <- read_parquet(here("data/ci_dyn/ci_rn.parquet"))
ci_rug.df <- read_parquet(here("data/ci_dyn/ci_rug.parquet"))
```

# process data for reaction norm
```{r}
# import labelling scheme
ez_label <- read.csv(here("data/ez_label.csv"))

# get si_label with ci cue range
si_ci_rug.df <- ci_rug.df %>% 
  mutate(label_si = case_when(
    label %in% c("I", "I1+I2") ~ "I",
    label %in% c("I log","I1+I2 log") ~ "I log",
    label %in% c("Ig", "Ig1+Ig2") ~ "Ig",
    label %in% c("Ig log") ~ "Ig log",
    label %in% c("sum", "I+Ig") ~ "I+Ig",
    label %in% c("sum log", "I+Ig log") ~ "I+Ig log",
    label == "R" ~ "R",
    label == "R log" ~ "R log",
    label %in% c("G", "G1+G2") ~ "G",
    label == "G log" ~ "G log"
  )) 

# get limit for si_rug
si_rug_lim.df <- si_rug.df %>% 
  filter(time <= 20) %>%
  group_by(label)%>% 
  summarise(min = min(value, na.rm = T)*0.9,
         max = max(value, na.rm = T)*1.1) %>% 
  select(label_si = label, min_si = min, max_si = max)

# filter to restriction conversion rate reaction norm range to cue ranges that appear in rug
## change to Inf/-inf to NA. Note that I am first joining with si rug lim to check which limit is larger, We will go with the cue range that has the largest span
ci_rug_lim.df <- si_ci_rug.df %>% 
  group_by(label) %>% 
  mutate(min = min(value, na.rm = T)*0.9,
         max = max(value, na.rm = T)*1.1) %>% 
  distinct(label, .keep_all = T) %>% 
  select(label, label_si, min, max)

rug_lim.final <- ci_rug_lim.df %>% left_join(si_rug_lim.df, by = "label_si") %>% 
  mutate(final_min = min(min, min_si),
         final_max = max(max, max_si))

# get second rug_lim.final for single infection
rug_lim.final2 <- rug_lim.final %>% 
  group_by(label_si) %>% 
  mutate(final_min = min(final_min, na.rm = T),
         final_max = max(final_max, na.rm = T))

# filter ci_rn by limit
ci_rn.df2 <- ci_rn.df %>% 
  left_join(rug_lim.final, by  = "label") %>% 
  group_by(label) %>% 
  filter(cue_range <= final_max & cue_range >= final_min)
```

# match single infection rn with coinfection 
```{r}
# get ci label to si rug and filter by limit
si_rn.df2 <- left_join(si_rn.df, select(ez_label, label_si, label_ci), by = c("label" = "label_si")) %>% 
  left_join(rug_lim.final, by  = c("label_ci" = "label")) %>% 
  group_by(label_ci) %>% 
  filter(cue_range <= final_max & cue_range >= final_min) %>% 
  select(cue_range, cr, label_ci, label_si)

# get ci label to si rug
si_rug.df2 <- select(si_rug.df, value, label_si = label) 
```


# plot reaction norm
```{r}
# join with ezlabel
ci_rn.df3 <- ci_rn.df2 %>% left_join(ez_label, by = c("label" = "label_ci"))
si_rn.df3 <- si_rn.df2 %>% left_join(ez_label, by = "label_ci")
ci_rug.df3 <- ci_rug.df %>% left_join(ez_label, by = c("label" = "label_ci"))
si_rug.df3 <- si_rug.df2 %>% left_join(ez_label, by = "label_si")

# redo order of cues
ci_rn.df3$label <- factor(ci_rn.df3$label, 
                              levels = c("I", "I log",
                                         "I1+I2", "I1+I2 log",
                                         "Ig", "Ig log",
                                         "I+Ig", "I+Ig log",
                                         "sum", "sum log",
                                         "G", "G log",
                                         "G1+G2", "Ig1+Ig2",
                                         "R", "R log")) 

si_rn.df3$label_ci <- factor(si_rn.df3$label_ci, 
                              levels = c("I", "I log",
                                         "I1+I2", "I1+I2 log",
                                         "Ig", "Ig log",
                                         "I+Ig", "I+Ig log",
                                         "sum", "sum log",
                                         "G", "G log",
                                         "G1+G2", "Ig1+Ig2",
                                         "R", "R log"))

ci_rug.df3$label <- factor(ci_rug.df3$label, 
                              levels = c("I", "I log",
                                         "I1+I2", "I1+I2 log",
                                         "Ig", "Ig log",
                                         "I+Ig", "I+Ig log",
                                         "sum", "sum log",
                                         "G", "G log",
                                         "G1+G2", "Ig1+Ig2",
                                         "R", "R log"))

si_rug.df3$label_ci <- factor(si_rug.df3$label_ci, 
                              levels = c("I", "I log",
                                         "I1+I2", "I1+I2 log",
                                         "Ig", "Ig log",
                                         "I+Ig", "I+Ig log",
                                         "sum", "sum log",
                                         "G", "G log",
                                         "G1+G2", "Ig1+Ig2",
                                         "R", "R log"))

# plot
opt_cue_pl.A <- ggplot() +
  geom_line(data = ci_rn.df3, aes(x = cue_range, y = cr, color = "Co-infection")) +
  geom_line(data = si_rn.df3, aes(x = cue_range, y = cr, color = "Single infection"), linetype = "dashed") +
  geom_rug(data = ci_rug.df3, aes(x = value), color = "#4575b4", sides = "t") +
  geom_rug(data = si_rug.df3, aes(x = value), color = "#fc8d59", sides = "b") +
  facet_wrap(~ez_label, scales = "free_x", ncol = 2) +
  theme_bw() +
  labs(y = "Conversion rate", x = "Cue range", color = "Infection") +
  scale_x_continuous(labels = function(x) format(x, scientific = T),
                     guide = guide_axis(check.overlap = TRUE)) + 
                    theme(axis.text.x = element_text(size = 7))  +
  scale_color_manual(values=c( "#4575b4", "#fc8d59"))

ggsave(here("figures/plos-bio/reaction_norm.png"), width = 7, height = 10)
```

#---------- plot best fitness --------------#
# import in data
```{r}
# single infection dynamics
si_dyn.df <- read_parquet(here("data/si_dyn/si_dyn_30.parquet")) 

# co-infection dynamics
ci_dyn.df <- read_parquet(here("data/ci_dyn/ci_dyn.parquet"))

ez_label <- read.csv(here("data/ez_label.csv"))
```

# process for final 20 days fitness
```{r}
# get single infection maximum tau_cum for 20 days
si_fitness.df <- si_dyn.df %>% 
  filter(variable == "tau_cum" & time == 20)

# get co-infection maximum tau_cum for 20 days
ci_fitness.df <- ci_dyn.df %>% 
  filter(variable == "tau_cum1" & time == 20)

# join together two dataframes and add labels
si_ci_fitness.df <- select(ci_fitness.df, fitness_ci = value, label_ci = label) %>% 
  left_join(ez_label, by = "label_ci") %>% 
  left_join(select(si_fitness.df, fitness_si = value, id_si = id), by = "id_si")
```

# plot
```{r}
opt_cue_pl.B <- ggplot() +
  geom_point(data = si_ci_fitness.df, aes(x = fitness_si, y = fitness_ci, color = ez_label_si, shape = ez_label_si), size = 2.5) +
  ggrepel::geom_label_repel(data = si_ci_fitness.df, aes(label = ez_label, x = fitness_si, y = fitness_ci),
                            fill = "white",xlim = c(-Inf, Inf), ylim = c(NA, NA)) +
  labs(x = "Optimal single-infection fitness", y = "Optimal Co-infection fitness (per strain)",
       color = "Single infection cue", shape = "Single infection cue") +
  scale_shape_manual(values = 15:24) +
  lims(x = c(8, 10.5)) +
  geom_vline(xintercept = 9.2, linetype = "dashed", alpha = 0.5) + # si good cue cutoff
  geom_hline(yintercept = 2.25, linetype = "dashed", alpha = 0.5) + # ci good cue cutoff
  theme_bw() +
  coord_cartesian(clip = "off") +
  theme(plot.margin=margin(r=0))
```

#-------------- timeseries conversion rate -----------------#
# process info for single infection
```{r}
# get fintess
si_cr.df <- si_dyn.df %>% 
  filter(time <= 20 & variable == "cr")

# good cue bad cue
si_cue.dv <- si_fitness.df %>% 
  mutate(classification = case_when(
    value > 9.2 ~ "High-performing",
    value <= 9.2 ~ "Poor-performing"
  ))

# join with classificaiton
si_cr.df2 <- si_cr.df %>% 
  left_join(select(si_cue.dv, id, classification, fitness_si = value), by = "id") %>% 
  left_join(ez_label, by = c("id" = "id_si"))

# split into top erforming and poor-performing cues
si_cr.high <- si_cr.df2 %>% filter(classification == "High-performing")
si_cr.poor <- si_cr.df2 %>% filter(classification == "Poor-performing")
si_cue.dv
```

# plot single infection conversion rate heatmap
```{r}
# plot poor performing
si_cr.pl1 <- ggplot() +
  geom_tile(data = si_cr.poor, aes(x = time, y = forcats::fct_reorder(ez_label_si, fitness_si), fill = value)) +
  labs(x = "Time (days)", y = "Low performing\nsingle infection cues", fill = "Conversion rate") +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlim(1, 20) +
  theme_bw()

# plot high perfomring
si_cr.pl2 <- ggplot() +
  geom_tile(data = si_cr.high, aes(x = time, y = forcats::fct_reorder(ez_label_si, fitness_si), fill = value)) +
  labs(x = "", y = "High performing\nsingle infection cues", fill = "Conversion rate") +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlim(1, 20) +
  theme_bw()
```

# co-infection conversion rate heatmap
# plot co-infeciton convesion rate heatmap
```{r}
# get fintess
ci_cr.df <- ci_dyn.df %>% 
  filter(time <= 20 & variable == "cr_1")

# good cue bad cue
ci_cue.dv <- ci_fitness.df %>% 
  mutate(classification = case_when(
    value > 2.25 ~ "High-performing",
    value <= 2.25 ~ "Poor-performing"
  ))

# join with classificaiton
ci_cr.df2 <- ci_cr.df %>% 
  left_join(select(ci_cue.dv, label, classification, fitness_ci = value), by = "label") %>% 
  left_join(ez_label, by = c("label" = "label_ci"))

# split into top erforming and poor-performing cues
ci_cr.high <- ci_cr.df2 %>% filter(classification == "High-performing")
ci_cr.poor <- ci_cr.df2 %>% filter(classification == "Poor-performing")
```

# plot
```{r}
# plot poor performing
ci_cr.pl1 <- ggplot() +
  geom_tile(data = ci_cr.poor, aes(x = time, y = forcats::fct_reorder(ez_label_si, fitness_ci), fill = value)) +
  labs(x = "Time (days)", y = "Low performing\nco-infection cues", fill = "Conversion rate") +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlim(1, 20) +
  theme_bw()

# plot high perfomring
ci_cr.pl2 <- ggplot() +
  geom_tile(data = ci_cr.high, aes(x = time, y = forcats::fct_reorder(ez_label_si, fitness_ci), fill = value)) +
  labs(x = "", y = "High performing\nco-infection cues", fill = "Conversion rate") +
  scale_fill_viridis_c(limits = c(0, 1)) +
  xlim(1, 20) +
  theme_bw()
```

#--------- assemble final figure --------------#
```{r}
# assemble the cr plots first
opt_cue_pl.C <- ggarrange(si_cr.pl2, ci_cr.pl2, si_cr.pl1, ci_cr.pl1, ncol = 2, nrow = 2, common.legend = T, align = "hv")

# assemble panel B and C
opt_cue_pl.BC <- ggarrange(opt_cue_pl.B, opt_cue_pl.C, ncol = 1, nrow = 2, labels = c("B", "C"), heights = c(0.75, 1))

# assemble final figure
ggarrange(opt_cue_pl.A, opt_cue_pl.BC, ncol = 2, nrow = 1, labels = c("A", ""), widths = c(0.75, 1))

# save figure
ggsave(here("figures/plos-bio/opt_cue.tiff"), units = "px", width = 2250, height = 1500, scale = 2, dpi=300, compression = "lzw")
```

#--------------------------------------------#
# dual cue optimization figure
#--------------------------------------------#
```{r}
source(here("functions/chabaudi_si_clean_high.R"))
source(here("functions/chabaudi_si_clean.R"))
source(here("functions/par_to_hm_te.R"))
```

#---------- plotting fitness of dual vs single cue opt ---------#
# import in previous data
```{r}
# dual cue fitness
dual_fitness.df <- read.csv(here("data/dual_cue_opt4/dual_cue_fitness_20.csv"))
## make label and filter out very low fitness
dual_fitness.df <- dual_fitness.df %>% 
  mutate(temp_label = gsub("log", "log10", label),
         temp_label_b = gsub("log", "log10", label_b),
         label_final = paste0(temp_label, "+", temp_label_b)) %>% 
  filter(value > 2)

# get single cue fitness
si_dyn.df <- read_parquet(here("data/si_dyn/si_dyn_30.parquet")) 
si_fitness.df <- si_dyn.df %>% 
  filter(variable == "tau_cum" & time == 20)

# join si and dual cue
dual_si_fitness.df <- dual_fitness.df %>% 
  left_join(select(si_fitness.df, id, si_fitness = value), by = "id") %>% 
  left_join(select(si_fitness.df, id_b = id, si_fitness_b = value), by = "id_b") %>% 
  mutate(si_fitness_max = ifelse(si_fitness > si_fitness_b, si_fitness, si_fitness_b),
         dual_label = gsub("log", "log10", paste(label, "+", label_b)))
```

# plot
```{r}
dual_si_fitness.pl <- ggplot() +
  geom_point(data = dual_si_fitness.df, aes(y = forcats::fct_reorder(dual_label, value), x = value, color = "Dual cue"), size = 2.5, alpha = 0.7) +
  geom_point(data = dual_si_fitness.df, aes(y = forcats::fct_reorder(dual_label, value), x = si_fitness_max, color= "Best single cue"), size = 2.5, alpha = 0.7) +
  labs(x = "Fitness", y = "Dual cue", color = "Cue") +
  scale_color_manual(values=c("#fc8d59", "#4575b4")) +
  geom_vline(xintercept = 9.883602) +
  geom_text(aes(x=9.883602, label="\nMaximum fitness", y = "G + R log10"), angle=90) +
  theme_bw()
```


#----------- time series conversion rate -------------#
# dynamics simulation of high parameter cues (these serve as reference points)
```{r}
# best dual cue combo
dual.cr <- chabaudi_si_clean(
  parameters_cr = c(4.446192033,	10.97518275,	1.38762817,	23.3059254,	-3.452052371,	-18.0070692,	39.66614226,	-3.545193141,	18.78350799),
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = seq(0, 20, by = 1e-3),
  cue_range =  seq(6, 7, by = 1/500),
  cue_range_b = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/500),
  cue = "R",
  cue_b = "I",
  log_cue = "log10",
  log_cue_b = "log10",
  solver = "vode",
  dyn = T
)

# when time is used as a cue (high parameter)
time.cr <- chabaudi_si_clean_high(
  parameters_cr = c(9.154314,  -7.570829, -22.506638 ,  3.382405 ,-13.453519 ,-17.011485  , 3.678181, -12.851895 ,-26.115158),
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = seq(0, 20, by = 1e-3),
  cue_range =  seq(0, 20, by = 1e-3),
  cue = "t",
  solver = "vode",
  dyn = T)

# when asexual iRBC is used as a cue (high flexibility)
I_high.cr <- chabaudi_si_clean_high(
  parameters_cr = c(1.296675,  3.544034 , 4.907484,  2.174249, -3.238309 ,-5.181614 ,-1.645072 , 1.834302 , 1.581011),
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = seq(0, 20, by = 1e-3),
  cue_range =  seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
  cue = "I",
  log_cue = "log10",
  solver = "vode",
  dyn = T)

# when asexual iRBC is used as a cue (normal flexibility)
I.cr <- chabaudi_si_clean(
  parameters_cr = c(5.463558,	2.383948,	-17.757281,	4.571835),
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = seq(0, 20, by = 1e-3),
  cue_range =  seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
  cue = "I",
  log_cue = "log10",
  solver = "vode",
  dyn = T)

# process 
I_high.cr2 <- I_high.cr %>% filter(variable == "cr") %>% mutate(label_new = "I log10 flexible") %>% select(-variable)

I.cr2 <- I.cr %>% filter(variable == "cr") %>% mutate(label_new = "I log10") %>% select(-variable)

time_high.cr2 <- time.cr %>% filter(variable == "cr") %>% mutate(label_new = "Time flexible") %>% select(-variable)

dual.cr2 <- dual.cr %>% filter(variable == "cr") %>% mutate(label_new = "R log10 + I log10") %>% select(-variable)

# combine
dual_cr.df <- rbind(I_high.cr2, I.cr2, time_high.cr2, dual.cr2)
```

# plot
```{r}
dual_cr.pl <- ggplot() +
  geom_line(data = dual_cr.df, aes(color = label_new, x = time, y = value), size = 1) +
  geom_point(data = dual_cr.df %>% filter(time%%1 == 0), aes(color = label_new, x = time, y = value, shape = label_new), size = 2) +
  labs(x = "Time (days)", y = "Conversion rate", color = "Cue", shape = "Cue") +
  xlim(0, 20) +
  scale_color_manual(values = c("#4575b4", "#91bfdb","#fc8d59","#fdcb44")) +
  theme_bw() +
  theme(legend.position="bottom") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE))
```

#------------ reaction norm heatmap of R log10 + I log10 ------------#
# process data
```{r}
# make heatmap df
R_I.hm <- par_to_hm_te(par = c(4.446192033,	10.97518275,	1.38762817,	23.3059254,	-3.452052371,	-18.0070692,	39.66614226,	-3.545193141,	18.78350799),
             cue_range = seq(6,	7, length.out = 500),
             cue_range_b = seq(0,	6.77815125, length.out = 500))

# process dynamics
R_I.dyn <- dual.cr %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value) %>% 
  mutate(log_R = log10(R),
         log_I = log10(I))
```

# plot
```{r}
dual_rn.pl <- ggplot() +
  geom_raster(data = R_I.hm, aes(x = cue_range_b, y = cue_range, fill = cr)) +
  scale_fill_viridis_c() +
  geom_path(data = R_I.dyn, aes(x = log_I, y = log_R), color = "white", arrow = arrow(angle = 30, length = unit(0.1, "inches"))) +
  geom_point(data = R_I.dyn %>% filter(row_number() %% 1000 == 1 & time <= 20), aes(x = log_I, y = log_R), color = "white") +
  xlim(0.99*min(hablar::s(R_I.dyn$log_I), na.rm = T), 1.01* max(hablar::s(R_I.dyn$log_I), na.rm = T)) +
  ylim(0.99*min(hablar::s(R_I.dyn$log_R), na.rm = T),1.01* max(hablar::s(R_I.dyn$log_R), na.rm = T)) +
  labs(y = "RBC log10", x = "Asexual iRBC log10", fill = "Conversion rate") +
  theme_dark()
```


#-------- assemble final figure -------------#
```{r}
# assemble panel B and C
dual_pl.BC <- ggarrange(dual_cr.pl, dual_rn.pl, align = "v", ncol = 1, labels = c("B", "C"))

# assemble panel A
dual.pl <- ggarrange(dual_si_fitness.pl, dual_pl.BC, ncol = 2, labels = c("A", ""), widths = c(1, 0.75))
ggsave(here("figures/plos-bio/dual_cue.tiff"), units = "px", width = 2250, height = 1400, scale = 1.5, dpi=300, compression = "lzw")
```

#-------------------------------------------#
# heterocue adoption 
#-------------------------------------------#

#------- static competition ---------------#
# calculate fitness difference for 20 days
```{r}
# get dynamics
static.ls <- list.files(path = here("data/ci_static/"), pattern = "*.parquet", full.names = T)
static.ls <- lapply(static.ls, read_parquet)

# get fitness at day 20 (optimized for 20 days)
static_fitness.ls <- mclapply(static.ls, function(x){
  x %>% filter(time == 20 & variable %in% c("tau_cum1", "tau_cum2"))
})
static_fitness.df <- do.call(rbind, static_fitness.ls)

static_fitness.df <- tidyr::pivot_wider(static_fitness.df, names_from = "variable", id_cols = c("id_1", "id_2")) %>% 
  group_by(id_1, id_2) %>% 
  mutate(fitness_difference = tau_cum1-tau_cum2)
write.csv(static_fitness.df, here("data/ci_static.csv"))
```

# import and process data
```{r}
# import in dataset
static.df <- read.csv(here("data/ci_static.csv"))
ez_label <- read.csv(here("data/ez_label.csv"))

# join with labelling
static.df2 <- static.df %>% 
  left_join(select(ez_label, id_ci, label_ci_1 = label_ci), by = c("id_1" = "id_ci")) %>% 
  left_join(select(ez_label, id_ci, label_ci_2 = label_ci), by = c("id_2" = "id_ci")) %>% 
  select(label_ci_1, label_ci_2, fitness_difference) %>% 
  mutate(label_ci_1 = gsub("log", "log10", label_ci_1),
         label_ci_2 = gsub("log", "log10", label_ci_2))

# get reverse order, which is simply invovles switching the cues around the multiplying the fitness by negative 1
static.df3 <- static.df2
names(static.df3) <- c("label_ci_2", "label_ci_1", "fitness_difference")
static.df3$fitness_difference <- static.df2$fitness_difference * -1

# join
static.df4 <- rbind(static.df2, static.df3)
# get mean
static.mean <- static.df4 %>% 
  group_by(label_ci_1) %>% 
  summarize(mean_fitness = mean(fitness_difference, na.rm = T))
```

# plot
```{r}
# heatmap
static.pl1 <- ggplot(data = static.df4, aes(x = label_ci_2, y = label_ci_1, fill = fitness_difference))+
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "#fc8d59", high = "#4575b4", mid = "white", 
   midpoint = 0, space = "Lab", lim = c(-0.95, 0.95), name="Fitness\ndifference") +
  theme_minimal() +  
  theme(legend.position = "left",
  axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  text = element_text(size = 15),
  axis.ticks = element_blank(),
  plot.margin=margin(r = 0)) + 
  labs(x = "Strain 2 cue", y = "Strain 1 cue") +
  coord_fixed()

# mean 
static.pl2 <- ggplot() +
  geom_bar(data =  static.mean, aes(y = label_ci_1, x = mean_fitness), stat = "identity") +
  labs(y = "", x = "Mean fitness\ndifference") +
  theme_bw() +
  theme(plot.margin = margin(l = 0),
       panel.grid.major = element_blank(),
  panel.background = element_blank(),
  text = element_text(size = 15),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank())

static.pl <- ggarrange(static.pl1, static.pl2, align = "hv", widths = c(1, 0.2))
ggsave(here("figures/plos-bio/static_heatmap.tiff"), width = 10, height = 6)
```

#------ invasion matrix --------------------#
# import in data (already 20 days )
```{r}
invade.df <- read.csv(here("data/ci_invasion.csv"))
```


# process data for invasion matrix
```{r}
invade.mat <- invade.df %>% 
  group_by(V1 = pmin(mut_id, res_id), V2 = pmax(mut_id, res_id)) %>% # group by cue competition, irregardless of order
  mutate(id_alt = paste0(V1, V2),
         invade = case_when(
           fitness > 0 ~ "invade",
           fitness < 0 ~ "not invade"
         )) %>% 
  group_by(id_alt) %>% 
  mutate(
    mut_is_V1 = case_when(
    mut_id == V1 ~ "V1_invade",
    mut_id != V1 ~ "V1_invaded")) %>% 
  arrange(id_alt) %>% 
  select(fitness, V1, V2, id_alt, invade, mut_is_V1) %>% 
  tidyr::pivot_wider(names_from = mut_is_V1, values_from = fitness) %>% 
  group_by(id_alt) %>% 
  mutate(V1_invade2 = gsub("NA", "", paste0(V1_invade, collapse = "")),
         V1_invaded2 = gsub("NA", "", paste0(V1_invaded, collapse = ""))) %>% 
  distinct(id_alt, .keep_all = T) %>% 
  mutate(
    category = case_when(
    V1_invade2 > 0 & V1_invaded2 > 0 ~ "Mutual invasion",
    V1_invade2 > 0 & V1_invaded2 < 0 ~ "Only strain 1 invasion",
    V1_invade2 < 0 & V1_invaded2 > 0 ~ "Only strain 2 invasion",
    V1_invade2 < 0 & V1_invaded2 < 0 ~ "Mutual non-invasion"
  )) %>% 
  select(V1, V2, invasion = category)

# for plotting, need to get all same cue vs same cue, which we will set to NA
invade.NA <- cbind.data.frame(`V1` = unique(invade.mat$V1),
      `V2` = unique(invade.mat$V1),
      invasion = NA)

invade.mat2 <- rbind(invade.mat, invade.NA)

# get label
invade.mat3 <- invade.mat2 %>% 
  left_join(select(ez_label, id_ci, V1_label = label_ci), by = c("V1" = "id_ci")) %>% 
  left_join(select(ez_label, id_ci, V2_label = label_ci), by = c("V2" = "id_ci")) %>% mutate(V1_label = gsub("log", "log10", V1_label),
                                      V2_label = gsub("log", "log10", V2_label))

# reorder so that ggplot do not mess up
library(gtools)

levels <- mixedsort(unique(c(invade.mat3$V1_label, invade.mat3$V2_label)))

invade.mat4 <- rbind(invade.mat3, 
                        invade.mat3 %>% 
                          rename(V2_label = V1_label, V1_label = V2_label) %>% 
                          select(V1_label, V2_label, invasion)) %>%
  mutate(across(ends_with("label"),
                ~factor(.x, levels = levels))) %>%
  filter(as.integer(V1_label) < as.integer(V2_label)) %>%
  mutate(V2_label = forcats::fct_rev(V2_label))
```

# plot invasion matrix
```{r}
# plot
invasion.pl1 <- ggplot(data = invade.mat4, aes(x = V2_label, y = V1_label, fill = invasion)) +
  geom_tile(color = "black") +
  theme_minimal() +  
  theme(legend.position = "right",
  axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  text = element_text(size = 15),
  axis.ticks = element_blank(),
  plot.margin=margin(r = 0)) + 
  labs(fill = "Invasion", x = "Strain 2 cue", y = "Strain 1 cue") +
  #scale_x_discrete(limits = rev) +
  scale_y_discrete(limits = rev) +
  scale_fill_manual(values = c("Only strain 1 invasion" = "#4575b4",
                               "Only strain 2 invasion" = "#fc8d59",
                               "Mutual invasion" = "#fee090")) +
  theme(legend.position = "none")
```

# create summary bar chart
```{r}
# create a stacked barchart for summary
## filter out na
invade.matalt <- invade.mat3 %>% na.exclude()

# get frquency from both sides. Note when grouping for V2, from the perspective of cue 2, scenarrio when strain 2 invade = strain 1 invade
invade.matalt1 <- invade.matalt %>% group_by(V1_label, invasion) %>% 
  summarize(frequency_1 = n())

invade.matalt2 <- invade.matalt %>%
  mutate(invasion_alt = case_when(
    invasion == "Only strain 1 invasion" ~ "Only strain 2 invasion",
    invasion == "Only strain 2 invasion" ~ "Only strain 1 invasion",
    invasion == "Mutual invasion" ~ "Mutual invasion",
    invasion == "Mutual non-invasion" ~ "Mutual non-invasion"
  )) %>% 
  group_by(V2_label, invasion_alt) %>% 
  summarize(frequency_2 = n())     

# full join and sum. has confirmed all of them add up to 14 
invade.matalt3 <- full_join(invade.matalt1, invade.matalt2, by = c("V1_label" = "V2_label", "invasion" = "invasion_alt"))

invade.matalt3[is.na(invade.matalt3)] <- 0
invade.matalt4 <- invade.matalt3 %>% 
  mutate(freq = frequency_1 + frequency_2) %>% 
  mutate(temp = case_when(
    invasion == "Only strain 1 invasion" ~ freq
  )) %>% 
  group_by(V1_label) %>% 
  mutate(invade_1_freq = max(temp, na.rm = T))
  
invasion.pl2 <- ggplot() +
  geom_bar(data = invade.matalt4, aes(x = freq, y = reorder(V1_label, invade_1_freq), fill = forcats::fct_rev(factor(invasion, levels = c("Only strain 1 invasion", "Only strain 2 invasion", "Mutual invasion", "Mutual non-invasion")))), stat = "identity") +
  labs(x = "Frequency", fill = "Invasion", y = "Strain 1 cue") +
  theme_bw()  +
  scale_fill_manual(values = c("Only strain 1 invasion" = "#4575b4",
                               "Only strain 2 invasion" = "#fc8d59",
                               "Mutual invasion" = "#fee090")) +
  theme(text = element_text(size = 15))
```

# plot together
```{r}
invasion.pl <- ggarrange(invasion.pl1, invasion.pl2, align = "h", common.legend = T, widths = c(2, 1))
ggsave(here("figures/plos-bio/invasion_heatmap.tiff"), width = 10, height = 6)
```

#------ effect cue perception -------#

# static
## logging
```{r}
# get non-logged pairings
static_nolog <- static.df2 %>% 
  mutate(cue_1 = trimws(gsub("\\ .*", "", label_ci_1)),
         log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none")) %>% 
  filter(log_1 == "none")


static_log <- static.df2 %>% 
  mutate(cue_1 = trimws(gsub("\\ .*", "", label_ci_1)),
         log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none")) %>% 
  filter(log_1 == "log")

static_log.df <- left_join(
  select(static_nolog, cue_1, label_ci_2, log_1, None = fitness_difference),
  select(static_log, cue_1, label_ci_2, log_1, Log = fitness_difference),
  by = c("cue_1", "label_ci_2")) %>% 
  filter(!is.na(None) & !is.na(Log)) %>% 
  mutate(classification = ifelse(Log > None, "Logged better", "Not logged better"))
```

# combined
```{r}
static_nocomb <- static.df2 %>% 
  mutate(cue_1 = ifelse(
    str_detect(label_ci_1, "sum"), "I+Ig",
    trimws(gsub("+1.*|\\ log", "", label_ci_1))
  ),
  log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none"),
    comb_1 = case_when(
    str_detect(label_ci_1, "1+|sum") ~ "comb",
    str_detect(label_ci_1, "1+|sum", negate = T) ~ "none" 
  )) %>% 
  filter(comb_1 == "none")

static_comb <- static.df2 %>% 
  mutate(cue_1 = ifelse(
    str_detect(label_ci_1, "sum"), "I+Ig",
    trimws(gsub("+1.*|\\ log", "", label_ci_1))
  ),
  log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none"),
    comb_1 = case_when(
    str_detect(label_ci_1, "1+|sum") ~ "comb",
    str_detect(label_ci_1, "1+|sum", negate = T) ~ "none" 
  )) %>% 
  filter(comb_1 == "comb")
  
static_comb.df <- left_join(
  select(static_nocomb, cue_1, label_ci_2, log_1, Self = fitness_difference),
  select(static_comb, cue_1, label_ci_2, log_1, Total = fitness_difference),
  by = c("cue_1", "log_1", "label_ci_2")) %>% 
  filter(!is.na(Total) & !is.na(Self)) %>% 
  mutate(classification = ifelse(Total > Self, "Total better", "Self better"))
```

# plot
```{r}
static_log.pl <- ggpaired(static_log.df, cond1 = "None", cond2 = "Log", line.color = "classification", alpha = 0.5) +
  labs(x = "Cue perception", y = "Fitness difference", color = "Effect") +
  scale_color_manual(values = c("Not logged better" = "#fc8d59", "Logged better" = "#4575b4"))

static_comb.pl <- ggpaired(static_comb.df, cond1 = "Total", cond2 = "Self", line.color = "classification", alpha = 0.5) +
  labs(x = "Cue perception", y = "Fitness difference", color = "Effect") +
  scale_color_manual(values = c("Total better" = "#fc8d59", "Self better" = "#4575b4"))

static_cue.pl <- ggarrange(static_log.pl, static_comb.pl, ncol = 2, align = "h")
ggsave(here("figures/plos-bio/static_cue_perception.tiff"), width = 7.5, height = 4)
```

# invasion
## proces data
```{r}
# join invade df with label because I am lazy
invade.df2 <- invade.df %>% 
  left_join(select(ez_label, id_ci, label_ci_1 = label_ci), by = c("mut_id" = "id_ci"))
```

# log
```{r}
# get non-logged pairings
invade_nolog <- invade.df2 %>% 
  mutate(cue_1 = trimws(gsub("\\ .*", "", label_ci_1)),
         log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none")) %>% 
  filter(log_1 == "none")


invade_log <- invade.df2 %>% 
  mutate(cue_1 = trimws(gsub("\\ .*", "", label_ci_1)),
         log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none")) %>% 
  filter(log_1 == "log")

invade_log.df <- left_join(
  select(invade_nolog, cue_1, res_id, log_1, None = fitness),
  select(invade_log, cue_1, res_id, log_1, Log = fitness),
  by = c("cue_1", "res_id")) %>% 
  filter(!is.na(None) & !is.na(Log)) %>% 
  mutate(classification = ifelse(Log > None, "Logged better", "Not logged better"))
```

# combined
```{r}
invade_nocomb <- invade.df2 %>% 
  mutate(cue_1 = ifelse(
    str_detect(label_ci_1, "sum"), "I+Ig",
    trimws(gsub("+1.*|\\ log", "", label_ci_1))
  ),
  log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none"),
    comb_1 = case_when(
    str_detect(label_ci_1, "1+|sum") ~ "comb",
    str_detect(label_ci_1, "1+|sum", negate = T) ~ "none" 
  )) %>% 
  filter(comb_1 == "none")

invade_comb <- invade.df2 %>% 
  mutate(cue_1 = ifelse(
    str_detect(label_ci_1, "sum"), "I+Ig",
    trimws(gsub("+1.*|\\ log", "", label_ci_1))
  ),
  log_1 = case_when(
           str_detect(label_ci_1, "log") ~ "log",
           str_detect(label_ci_1, "log", negate = T) ~ "none"),
    comb_1 = case_when(
    str_detect(label_ci_1, "1+|sum") ~ "comb",
    str_detect(label_ci_1, "1+|sum", negate = T) ~ "none" 
  )) %>% 
  filter(comb_1 == "comb")
  
invade_comb.df <- left_join(
  select(invade_nocomb, cue_1, res_id, log_1, Self = fitness),
  select(invade_comb, cue_1, res_id, log_1, Total = fitness),
  by = c("cue_1", "log_1", "res_id")) %>% 
  filter(!is.na(Total) & !is.na(Self)) %>% 
  mutate(classification = ifelse(Total > Self, "Total better", "Self better"))
```

# plot
```{r}
invade_log.pl <- ggpaired(invade_log.df, cond1 = "None", cond2 = "Log", line.color = "classification", alpha = 0.5) +
  labs(x = "Cue perception", y = "Fitness difference", color = "Effect") +
  scale_color_manual(values = c("Not logged better" = "#fc8d59", "Logged better" = "#4575b4"))

invade_comb.pl <- ggpaired(invade_comb.df, cond1 = "Total", cond2 = "Self", line.color = "classification", alpha = 0.5) +
  labs(x = "Cue perception", y = "Fitness difference", color = "Effect") +
  scale_color_manual(values = c("Total better" = "#fc8d59", "Self better" = "#4575b4"))

invade_cue.pl <- ggarrange(invade_log.pl, invade_comb.pl, align = "h", ncol = 2)
ggsave(here("figures/plos-bio/invasion_cue_perception.tiff"), width = 7.5, height = 4)
```

#-------------------------------------------#
# Partitioning best cue
#-------------------------------------------#
#------- single infection -----------#
```{r}

```

#------- dual cue ------------------#
```{r}

```


#-------------------------------------------#
Best cue adoption consequences
#-------------------------------------------#
# get data for disease curves
```{r}
# single infection dynamics
si_dyn.df <- read_parquet(here("data/si_dyn/si_dyn_30.parquet")) 

# co-infection dynamics
ci_dyn.df <- read_parquet(here("data/ci_dyn/ci_dyn.parquet"))

# dual cue dynamics
dual_dyn.df <- read_parquet(here("data/dual_cue_dyn/dual_cue_dyn.parquet"))
```





