---
title: "R Notebook"
output: html_notebook
---
```{r}
library(dplyr)
library(ggplot2)
library(here)
library(deSolve)
library(crone)
library(optimParallel)
library(doParallel)
library(doRNG)
library(arrow)
library(stringr)
library(parallel)
library(ggpubr)
```

```{r}
source(here("functions/chabaudi_si_clean.R"))
```

Resolving single infection optimization
Previous attempts at optimization resulted in an infection curve that went unresolved. Here, I cannot get the infection period to resolve. Increasing the infection period causes the infection dynamic of the parasite to become flattened. 
```{r}
time_range <- seq(0, 30, by = 1e-3)
```

```{r}
source(here("functions/par_to_df.R"))
source(here("functions/chabaudi_si_clean.R"))

parameters_tsukushi <- c(R1 = 8.89*10^6, # slightly higher
                lambda = 3.7*10^5,
                mu = 0.025, 
                p = 8*10^-6, # doubled form original
                alpha = 1, 
                alphag = 2, 
                beta = 5.721, 
                mum = 48, 
                mug = 4, 
                I0 = 43.85965, 
                Ig0 = 0, 
                a = 150, 
                b = 100, 
                sp = 1,
                psin = 16.69234,
                psiw = 0.8431785,
                phin = 0.03520591, 
                phiw = 550.842,
                iota = 2.18*(10^6),
                rho = 0.2627156)

```

```{r}
opt_local <- function(
  par, # initial parameter sets
  cue, #
  log,
  cue_range,
  dyn = F
  ){
  # start cluster
  cl <- makeCluster(detectCores()); setDefaultCluster(cl = cl)
  
  # start optimization LFBGS
  res <- optimParallel(
    par = par,
    fn = chabaudi_si_clean, 
    control = list(trace = 6, fnscale = -1),
    immunity = "tsukushi",
    parameters = parameters_tsukushi,
    time_range = time_range,
    cue = cue,
    log_cue = log,
    cue_range = cue_range,
    solver = "vode")
  
  # close cluster
  stopCluster(cl)
  
  return (res)
}
```

# I non-logged
```{r}
I_loc.op <- opt_local(par = rep(0.5, 4),
                            cue = "I",
                            log = "none",
                            cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000))

I_loc_20.dyn <- chabaudi_si_clean(parameters_cr = c(0.1135384,  -39.0585154, 1941.9082813,  119.8228980),
                               parameters= parameters_tsukushi,
                               immunity = "tsukushi",
                               time_range = seq(0,25,0.001),
                               cue = "I",
                               log_cue = "none",
                               cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                               dyn = T)

I_loc_25.dyn <- chabaudi_si_clean(parameters_cr = c(-0.4002295,  3.2517089,  0.6089734,  0.4916482),
                               parameters= parameters_tsukushi,
                               immunity = "tsukushi",
                               time_range = seq(0,25,0.001),
                               cue = "I",
                               log_cue = "none",
                               cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                               dyn = T)

I_loc_30.dyn <- chabaudi_si_clean(parameters_cr = c(-0.4632307,  -12.7384555, 1898.9561735, -292.2138265),
                               parameters= parameters_tsukushi,
                               immunity = "tsukushi",
                               time_range = seq(0,30,0.001),
                               cue = "I",
                               log_cue = "none",
                               cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                               dyn = T)

# 20 days optimziation
#0.1135384  -39.0585154 1941.9082813  119.8228980
#8.49777

# 25 days 
#-0.4002295  3.2517089  0.6089734  0.4916482
#9.860488

# 30 days
#-0.4632307  -12.7384555 1898.9561735 -292.2138265
#11.81462

# plotting together, we cannot get the infeciton cycle to resolve. By elongating the infection period, we see a flattening of parasite dynamics that prolongs the infection
ggplot()+ 
  geom_line(data = I_loc_30.dyn, aes(x = time, y = value, color = "30")) +
  geom_line(data = I_loc_25.dyn, aes(x = time, y = value, color = "25")) +
  geom_line(data = I_loc_20.dyn, aes(x = time, y = value, color = "20")) +
  facet_wrap(~variable, scales = "free") +
  theme_bw()
```

# I logged
Still no full resolution even with 30 days, although it is closer to resolution than 20 days optimization. Extending the simulation to 25 days allows the optimum 20 days to fully resolve. Hence, for future purposes, it might be better to optimize based on 20 days for optimal acute infection. However, for simulation go to 30 days to ensure fulll resolution 
```{r}
I_log_loc.op <- opt_local(par = rep(0.5, 4),
                            cue = "I",
                            log = "log10",
                            cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000))

# 20 days
#5.463558   2.383948 -17.757281   4.571835
#9.494991

# 30 days
# 3.53672545  -0.03564319 -12.38102999   5.20523629
# 13.27947
I_log_loc_20.dyn <- chabaudi_si_clean(parameters_cr = c(5.463558,   2.383948 ,-17.757281 ,  4.571835),
                               parameters= parameters_tsukushi,
                               immunity = "tsukushi",
                               time_range = seq(0,30,0.001),
                               cue = "I",
                               log_cue = "log10",
                               cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
                               dyn = T)

I_log_loc_30.dyn <- chabaudi_si_clean(parameters_cr = c(3.53672545,  -0.03564319, -12.38102999,   5.20523629),
                               parameters= parameters_tsukushi,
                               immunity = "tsukushi",
                               time_range = seq(0,30,0.001),
                               cue = "I",
                               log_cue = "log10",
                               cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
                               dyn = T)

ggplot()+ 
  geom_line(data = I_log_loc_30.dyn, aes(x = time, y = value, color = "30")) +
  geom_line(data = I_log_loc_20.dyn, aes(x = time, y = value, color = "20")) +
  facet_wrap(~variable, scales = "free") +
  theme_bw()
```


#---------------------------------------------------------#
# Co-infection optimization
#---------------------------------------------------------#
```{r}
source(here("functions/par_to_df.R"))
source(here("functions/chabaudi_si_clean.R"))
source(here("functions/chabaudi_ci_clean.R"))
parameters_tsukushi <- c(R1 = 8.89*10^6, # slightly higher
                lambda = 3.7*10^5,
                mu = 0.025, 
                p = 8*10^-6, # doubled form original
                alpha = 1, 
                alphag = 2, 
                beta = 5.721, 
                mum = 48, 
                mug = 4, 
                I0 = 43.85965, 
                Ig0 = 0, 
                a = 150, 
                b = 100, 
                sp = 1,
                psin = 16.69234,
                psiw = 0.8431785,
                phin = 0.03520591, 
                phiw = 550.842,
                iota = 2.18*(10^6),
                rho = 0.2627156)

time_range <- seq(0, 20, by = 1e-3)
```

#---------------------------------#
optimizing co-infection
#---------------------------------#

fully static run with the optimal parameter for each cue competing against each other. parasites are not allowed to alter their strategy. In biology terms, think of when intrastrain competition>>interstrain competition, thereby maintaining the original optimizal parameters.

```{r}
source(here("functions/chabaudi_ci_clean.R"))
```

```{r}


ci_comp_static <- function(df){
  
  
  
}
```

# create cue_range df
if _i at end of cue, suggesting of I1/I2 (self-differentation possible).
```{r}
# code. 
cue_range.df <- cbind.data.frame(
  cue = c("I_i", "I_i",
          "I1+I2", "I1+I2",
          "Ig_i", "Ig_i",
          "Ig1+Ig2", "Ig1+Ig2",
          "sum", "sum",
          "G_i", "G_i"),
  log = c("none", "log",
          "none", "log",
          "none", "log",
          "none", "log",
          "none", "log",
          "none", "log"),
  low = c(0, 0,
          0, 0,
          0, 0,
          0, 0,
          0, 9),
  high = c(6*(10^6), log10(6*(10^6)),
           6*(10^6), log10(6*(10^6)),
           6*(10^6), log10(6*(10^6)),
           6*(10^6), log10(6*(10^6)),
           6*(10^6), log10(6*(10^6))),
  by = c((6*(10^6))/5000, (log10(6*(10^6)))/5000,
         (6*(10^6))/5000, (log10(6*(10^6)))/5000,
         (6*(10^6))/5000, (log10(6*(10^6)))/5000,
         (6*(10^6))/5000, (log10(6*(10^6)))/5000,
         (6*(10^6))/5000, (log10(6*(10^6)))/5000)
)
```


```{r}
df <- cbind.data.frame(
  cue = c("I", "Ig", "sum"),
  log = c("none", "log", "log"),
  var1 = c(1.931059, 35.6294, 435.1835),
  var2 = c(-63.786820, -65.6898, -565.3985),
  var3 = c(1295.819775, -11.5609, -395.7242),
  var4 = c(-31.580440, -52.0779, -446.5950)
) %>% mutate(id = paste0(cue, "_", log))

df2 <- df %>% 
  setNames(paste0(names(df), '_2')) %>% 
  tidyr::crossing(df, .name_repair = "check_unique") %>% 
  janitor::clean_names()

df3 <- df2[!duplicated(apply(df2, 1, function(x) paste(sort(x),collapse=''))),]


df3[1,]$cue
chabaudi_ci_clean(parameters_cr_1 = c(df3[1,]$var1, df3[1,]$var2, df3[1,]$var3, df3[1,]$var4),
                  parameters_cr_2 = c(df3[1,]$var1_2, df3[1,]$var2_2, df3[1,]$var3_2, df3[1,]$var4_2),
                  immunity = "tsukushi",
                  parameters = parameters_tsukushi,
                  cue_1 = df3[1,]$cue,
                  cue_2 = df3[1,]$cue_2,
                  cue_range_1 = ,
                  cue_range_2 = ,
                  log_cue_1 = ,
                  log_cue_2 = ,
                  solver = "vode")
```


