---
title: "report16"
output: html_document
---
```{r}
library(dplyr)
library(ggplot2)
library(here)
library(deSolve)
library(crone)
library(optimParallel)
library(doParallel)
library(doRNG)
library(arrow)
library(stringr)
library(parallel)
library(ggpubr)
```

# analyzing MC partitioning results

# import in all fitness files
```{r}
file_ls <- list.files(path = here("data/MC_partitioned/"), pattern = "*.csv", full.names = T)
name_ls <- list.files(path = here("data/MC_partitioned/"), pattern = "*.csv")
name_ls <- gsub("*.csv", "", name_ls)

# 60, which is about right
length(file_ls)

# read in files
fitness.ls <- lapply(file_ls, read.csv)

# assign unique ID
fitness.ls <- mapply(cbind, fitness.ls, "ID" = name_ls, SIMPLIFY = F)
```

# process data
```{r}
# get metainfo from ID
fitness.ls2 <- mclapply(fitness.ls, function(x){
  id_col <- x$ID
  # string split to extract all info
  cue <- unlist(str_split(unique(id_col), pattern = "_"))[[3]]
  log <- unlist(str_split(unique(id_col), pattern = "_"))[[4]]
  rand_var <- unlist(str_split(unique(id_col), pattern = "_"))[[5]]
  
  # get mean
  mean_fitness <- mean(x$max_fitness)
  # get sd
  sd_fitness <- sd(x$max_fitness)
  
  # bind results
  res <- cbind(x, cue= cue, log = log, rand_var = rand_var, mean_fitness = mean_fitness, sd_fitness = sd_fitness)
  return(res)
})
```

# Get reference data
```{r}
reference_ls <- list.files(path = here("data/MC2"), pattern = "*.csv", full.names = T)
reference_name.ls <- gsub("*.csv", "", list.files(path = here("data/MC2/"), pattern = "*.csv"))

# read in the files
reference.ls <- lapply(reference_ls, read.csv)

# assign unique ID
reference.ls <- mapply(cbind, reference.ls, "ID" = reference_name.ls, SIMPLIFY = F)

# get meta data
reference.ls2 <- mclapply(reference.ls, function(x){
  id_col <- x$ID
  # string split to extract all info
  cue <- unlist(str_split(unique(id_col), pattern = "_"))[[2]]
  # get log
  third_col <- unlist(str_split(unique(id_col), pattern = "_"))[[3]]
  log <- ifelse(third_col == "log", "log10", "none")
  
  # get mean
  mean_fitness <- mean(x$max_fitness)
  
  # get sd
  sd_fitness <- sd(x$max_fitness)
  
  # bind results
  res <- cbind(x, cue= cue, log = log, rand_var = "all", ref_mean_fitness = mean_fitness, ref_sd_fitness = sd_fitness)
  return(res)
})
```

# combine MC partitioned and reference df
```{r}
# get unique column values for each cue, log, and rand_var combo
fitness.ls3 <- do.call(rbind, fitness.ls2)
fitness.ls3 <- fitness.ls3 %>% dplyr::distinct(ID, .keep_all = T)

# repeat with reference
reference.ls3 <- do.call(rbind, reference.ls2)
reference.ls3 <- reference.ls3 %>% dplyr::distinct(ID, .keep_all = T)

# combine!
ref_fit.df <- left_join(fitness.ls3, reference.ls3, by = c("cue" = "cue", "log"= "log"))
```

# compute proportion fitness and variation
```{r}
ref_fit.df2 <- ref_fit.df %>% 
  mutate(p_sd = sd_fitness/ref_sd_fitness,
         p_mean = ref_mean_fitness/mean_fitness,
         cue_log = paste0(cue, "_", log))
```

# plot!
```{r}
# variation
fig1a <- ggplot() +
  geom_tile(data = ref_fit.df2 , aes(x = cue, y = `rand_var.x`, fill = p_sd)) +
  facet_wrap(~log) +
  theme_bw() +
  viridis::scale_fill_viridis() +
  labs(x = "Cue", y = "Variable randomized", fill = expression(frac(sd("1 var randomized"), sd("all var randomized")))) +
  theme(legend.position="top")

# mean fitness
fig1b <- ggplot() +
  geom_tile(data = ref_fit.df2 , aes(x = cue, y = `rand_var.x`, fill = p_mean)) +
  facet_wrap(~log) +
  theme_bw() +
  viridis::scale_fill_viridis() +
  labs(x = "Cue", y = "Variable randomized", fill = expression(frac(Mean("all var randomized"), Mean("1 var randomized")))) +
  theme(legend.position="top")

mc_partition <- ggarrange(fig1a, fig1b, ncol = 1)

ggsave(here("figures/report16/mc_partition.png"), height = 10)
```

# create new MC fitnes
# read MC data
```{r}
# read in dymamics
mc_G_log.dyn <- read_parquet(here("data/MC2/mc_G_log_dyn.parquet"))
mc_G.dyn <- read_parquet(here("data/MC2/mc_G_dyn.parquet"))
mc_R_log.dyn <- read_parquet(here("data/MC2/mc_R_log_dyn.parquet"))
mc_R.dyn <- read_parquet(here("data/MC2/mc_R_dyn.parquet"))
mc_I_log.dyn <- read_parquet(here("data/MC2/mc_I_log_dyn.parquet"))
mc_I.dyn <- read_parquet(here("data/MC2/mc_I_dyn.parquet"))
mc_Ig_log.dyn <- read_parquet(here("data/MC2/mc_Ig_log_dyn.parquet"))
mc_Ig.dyn <- read_parquet(here("data/MC2/mc_Ig_dyn.parquet"))
mc_I_Ig_log.dyn <- read_parquet(here("data/MC2/mc_I+Ig_log_dyn.parquet"))
mc_I_Ig.dyn <- read_parquet(here("data/MC2/mc_I+Ig_dyn.parquet"))

# read in fitness
mc_G_log.fitness <- read.csv(here("data/MC2/mc_G_log_fitness.csv"))
mc_G.fitness <- read.csv(here("data/MC2/mc_G_fitness.csv"))
mc_R_log.fitness <- read.csv(here("data/MC2/mc_R_log_fitness.csv"))
mc_R.fitness <- read.csv(here("data/MC2/mc_R_fitness.csv"))
mc_I_log.fitness <- read.csv(here("data/MC2/mc_I_log_fitness.csv"))
mc_I.fitness <- read.csv(here("data/MC2/mc_I_fitness.csv"))
mc_Ig_log.fitness <- read.csv(here("data/MC2/mc_Ig_log_fitness.csv"))
mc_Ig.fitness <- read.csv(here("data/MC2/mc_Ig_fitness.csv"))
mc_I_Ig_log.fitness <- read.csv(here("data/MC2/mc_I+Ig_log_fitness.csv"))
mc_I_Ig.fitness <- read.csv(here("data/MC2/mc_I+Ig_fitness.csv"))
```

# examine variation
```{r}
# plot fitness vs iteration
fitness.df <- rbind(
  cbind(mc_G_log.fitness, id = "Gametocyte\nlog10"),
  cbind(mc_G.fitness, id = "Gametocyte"),
  cbind(mc_R_log.fitness, id = "RBC log10"),
  cbind(mc_R.fitness, id = "RBC"),
  cbind(mc_I_log.fitness, id = "Asexual iRBC\nlog10"),
  cbind(mc_I.fitness, id = "Asexual iRBC"),
  cbind(mc_Ig_log.fitness, id = "Sexual iRBC\nlog10"),
  cbind(mc_Ig.fitness, id = "Sexual iRBC"),
  cbind(mc_I_Ig_log.fitness, id = "Total iRBC\nlog10"),
  cbind(mc_I_Ig.fitness, id = "Total iRBC")
)

# fitness vs time
#fig2a <- ggplot() +
#  geom_line(data = fitness.df, aes(x = X, y = max_fitness), size =0.5) +
#  facet_wrap(~id, ncol = 2) +
#  labs(x = "Iteration", y = "Maximum fitness", color = "Cue") +
#  theme_bw() +
#  guides(color = FALSE, fill = F)

# quantify variance and mean
fitness_var.df <- fitness.df %>% 
  dplyr::group_by(id) %>% 
  dplyr::summarise(std = sd(max_fitness),
                   mean = mean(max_fitness)) %>% 
  dplyr::mutate(id = forcats::fct_reorder(factor(id), std, .desc = F))

# original rank by standard deviation
##fig2b <- ggplot() +
##  geom_col(data = fitness_var.df, aes(y = id, x = std)) +
  #scale_fill_manual(values = c("#619CFF", "#B79F00", "#F564E3", "#00BA38", "#F8766D", "#00BFC4")) +
##  theme_bw() +
#  labs(y = "Cue", x = "Standard deviation") +
#  guides(fill = F, color = F)

fig2a <- ggplot() +
  geom_point(data = fitness_var.df, aes(x = mean, y = std)) +
  ggrepel::geom_label_repel(data = fitness_var.df, aes(x = mean, y = std, label = id)) +
  theme_bw() +
  labs(x = "Mean fitness", y = "Fitness standard deviation")

fitness_var.df
```

# plot violin with difference in deterministic model fitness and mean model fitness
```{r}
# get deterministic df
det.df <- data.frame(id = fitness_var.df$id, fitness =  c(8.49777, 9.494991, 8.58856, 9.561373, 8.23991, 8.181604, 8.569285, 9.618812, 8.854682, 9.573291))

fig2b <- ggplot() +
  geom_violin(data = fitness.df, aes(y = id, x = max_fitness), alpha = 0.5) +
  geom_point(data = fitness_var.df, aes(y = id, x = mean), color = "red") +
  geom_point(data = det.df, aes(y = id, x = fitness), shape = 17) +
  labs(x = "Fitness", y = "Cue") +
  theme_bw()

fig2b
```

# get deviation from ideal cr
```{r}
source(here("functions/chabaudi_si_clean.R"))
dev_cr <- function(dyn, par, cue, cue_range, log){
  # get reference dynamics
  ref_dyn <- chabaudi_si_clean(
              parameters_cr = par,
              parameters = parameters_tsukushi, 
              time_range = time_range, 
              cue = cue, 
              cue_range = cue_range, 
              log_cue = log,
              immunity = "tsukushi",
              solver = "vode",
              dyn = TRUE)
  
  # convert to wide
  ref_dyn.wide <- tidyr::pivot_wider(ref_dyn, names_from = variable, values_from = value)
  
  # get 95% CI. Use t score of 1.960439 given our sample size of 5000. Margin of error = t*(sd/sqrt(n))
  mc_sum <- dyn %>%
  dplyr::group_by(time) %>% 
  dplyr::summarise(cr_top = mean(cr)+sd(cr),
                   cr_bot = mean(cr)-sd(cr),
                   tau_top = mean(tau)+sd(tau),
                   tau_bot = mean(tau)-sd(tau))
  
  # return list
  return(list(ref_dyn.wide, mc_sum))
}
```

# run tests
```{r}
# G log
G_log.diff <- dev_cr(mc_G_log.dyn, 
                    par = c(1.211521, -3.936778,-1.312944, -1.285713),
                    cue = "G",
                    cue_range = seq(0, log10(6*(10^4)), (log10(6*(10^4)))/5000),
                    log= "log10")

# G
G.diff <- dev_cr(mc_G.dyn, 
                 par = c(0.04061288,   -9.31445958,   74.13015506, -431.59843638),
                    cue = "G",
                    cue_range = seq(0, 6*(10^4), (6*(10^4))/5000),
                    log= "none")

# R log
R_log.diff <- dev_cr(mc_R_log.dyn, 
                     par = c(48.70772, 105.18882, -81.80625, -44.60051),
                    cue = "R",
                    cue_range = seq(log10(10^6), log10(10^7), by = (log10(10^7)-log10(10^6))/5000),
                    log= "log10")

# R
R.diff <- dev_cr(mc_R.dyn, 
                 par = c(7.0976026,  30.1295151, -28.0565613,  -0.2788713),
                    cue = "R",
                    cue_range = seq((10^6), (10^7), by = ((10^7)-(10^6))/5000),
                    log= "none")

# I log
I_log.diff <- dev_cr(mc_I_log.dyn, 
                     par = c(5.463558,   2.383948, -17.757281,   4.571835),
                    cue = "I",
                    cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
                    log= "log10")

# I
I.diff <- dev_cr(mc_I.dyn, 
                     par = c(0.1135384,  -39.0585154, 1941.9082813,  119.8228980),
                    cue = "I",
                    cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                    log= "none")

# Ig log
Ig_log.diff <- dev_cr(mc_Ig_log.dyn, 
                  par = c(1.74633, -1.78582, -5.52834, 1.57344 ),
                    cue = "Ig",
                    cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
                    log= "log10")

# Ig
Ig.diff <- dev_cr(mc_Ig.dyn, 
                  par = c(0.05234791,  -46.03368629, 1897.61313455,   70.57689145),
                    cue = "Ig",
                    cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                    log= "none")


# I+Ig log
I_Ig_log.diff <- dev_cr(mc_I_Ig_log.dyn, 
                    par = c(3.594042,   4.157744, -13.530672,   2.599905),
                    cue = "I+Ig",
                    cue_range = seq(0, log10(6*(10^6)), by = (log10(6*(10^6)))/5000),
                    log= "log10")


# I+Ig
I_Ig.diff <- dev_cr(mc_I_Ig.dyn, 
                    par = c(0.3159297,  -46.1104558, 1250.7529081,   -6.1982093),
                    cue = "I+Ig",
                    cue_range = seq(0, 6*(10^6), by = (6*(10^6))/5000),
                    log= "none")

```

# plot
```{r}
# cr time series
## concatenate data
reference.df <- rbind(
  cbind(G_log.diff[[1]], cue = "Gametocyte log10"),
  cbind(G.diff[[1]], cue = "Gametocyte"),
  cbind(R_log.diff[[1]], cue = "RBC log10"),
  cbind(R.diff[[1]], cue = "RBC"),
  cbind(I_log.diff[[1]], cue = "Asexual iRBC log10"),
  cbind(I.diff[[1]], cue = "Asexual iRBC"),
  cbind(Ig_log.diff[[1]], cue = "Sexual iRBC log10"),
  cbind(Ig.diff[[1]], cue = "Sexual iRBC"),
  cbind(I_Ig_log.diff[[1]], cue = "Total iRBC log10"),
  cbind(I_Ig.diff[[1]], cue = "Total iRBC")
)

## different data contenate
diff.df <- rbind(
  cbind(G_log.diff[[2]], cue = "Gametocyte log10"),
  cbind(G.diff[[2]], cue = "Gametocyte"),
  cbind(R_log.diff[[2]], cue = "RBC log10"),
  cbind(R.diff[[2]], cue = "RBC"),
  cbind(I_log.diff[[2]], cue = "Asexual iRBC log10"),
  cbind(I.diff[[2]], cue = "Asexual iRBC"),
  cbind(Ig_log.diff[[2]], cue = "Sexual iRBC log10"),
  cbind(Ig.diff[[2]], cue = "Sexual iRBC"),
  cbind(I_Ig_log.diff[[2]], cue = "Total iRBC log10"),
  cbind(I_Ig.diff[[2]], cue = "Total iRBC")
)


# plot conversion time series
fig2c <- ggplot() +
  geom_line(data = reference.df, aes(x = time, y = cr)) +
  geom_ribbon(data = diff.df, aes(x = time, ymin = cr_bot, ymax = cr_top), alpha = 0.3, fill = "red") +
  facet_wrap(~cue, ncol = 2) +
  labs(x = "Time (days)", y = "Conversion rate") +
  theme_bw()
  
# plot fitness timeseries. When if tiness lost? At the latter part
fig2d <- ggplot() +
  geom_line(data = reference.df, aes(x = time, y = tau)) +
  geom_ribbon(data = diff.df, aes(x = time, ymin = tau_bot, ymax = tau_top), alpha = 0.3, fill = "red") +
  facet_wrap(~cue, ncol = 2) +
  labs(x = "Time (days)", y = "Transmission potential") +
  theme_bw()

```

# plot together
```{r}
fig1ab <- ggarrange(fig2b,mc_partition, labels = c("A", "B"))

fig1cd <- ggarrange(fig2c, fig2d, labels = c("C", "D"))

fig1 <- ggarrange(fig1ab, fig1cd, ncol = 1, heights = c(1.5,1))
ggsave(here("figures/report16/MC_final.png"), height = 10, width = 10)
```

