---
title: "report5"
output: html_document
---

# Load libraries
```{r}
library(deSolve)
library(splines)
library(optimParallel)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(magrittr)
library(Rcpp)
library(microbenchmark)
library(GA)
library(here)
```

# Load functions
```{r}
source(here("functions/chabaudi_ci_opt_lag.R"))
source(here("functions/co_infection_opt.R"))
source(here("functions/par_to_df.R"))
source(here("functions/chabaudi_si_opt_lag.R"))
source(here("functions/total_parasite.R"))
```

# define parameters
```{r}
parameters <- c(R1 = 8.5*10^6,
                lambda = 3.7*10^5,
                mu = 0.025, 
                p = 4*10^-6,
                alpha = 1, 
                alphag = 2, 
                beta = 10, 
                mum = 48, 
                mug = 4, 
                I0 = 43.85965, 
                Ig0 = 0, 
                a = 150, 
                b = 100, 
                sp = 1)

parameters_tsukushi <- c(R1 = 8.89*10^6, # slightly higher
                lambda = 3.7*10^5,
                mu = 0.025, 
                p = 8*10^-6, # doubled form original
                alpha = 1, 
                alphag = 2, 
                beta = 5.721, 
                mum = 48, 
                mug = 4, 
                I0 = 43.85965, 
                Ig0 = 0, 
                a = 150, 
                b = 100, 
                sp = 1,
                psin = 16.69234,
                psiw = 0.8431785,
                phin = 0.03520591, 
                phiw = 550.842,
                iota = 2.18*(10^6),
                rho = 0.2627156)

time_range <- seq(0, 20, by = 1e-3)
I_range <- seq(0, 2.18*10^6, by = 1000) 
I_range_large <- seq(0, (3.7*10^5)*(8.5*10^6)/((3.7*10^5)-0.025*(8.5*10^6)), by = 1000) 
I_range_mid <- seq(0, 5*10^6, by = 1000)
I_range_mid2 <- seq(0, 4*10^6, by = 1000)
I_range_mid4 <- seq(0, 6*10^6, by = (6*10^6)/5000)

# logged counterpart
I_range_mid2_log <- seq(0, log(4*10^6), by = log(4*10^6)/5000)
I_range_mid3_log <- seq(0, log(5*10^6), by = log(5*10^6)/5000)
I_range_mid4_log <- seq(0, log(6*10^6), by = log(6*10^6)/5000)
I_range_mid4_log10 <- seq(0, log10(6*10^6), by = log10(6*10^6)/5000)
I_range_large_log <- seq(0, log((3.7*10^5)*(8.5*10^6)/((3.7*10^5)-0.025*(8.5*10^6))), by = log((3.7*10^5)*(8.5*10^6)/((3.7*10^5)-0.025*(8.5*10^6)))/10000)


# R_range
R_range_log10 <- seq(0, log10(8.89*10^6), by = log10(8.89*10^6)/10000)
```

#-----------------#
Exploring starting doses dynamics
#------------------#

#10^4 of one strain and 10^5 of residence strain
# oops. Some errors with previous model. Previous has starting density of 10^2 parasites each.


# testing single strain lagged model as proof of concept. What we did: for all conditional equation, add delay. Delayed the pulse beta function
```{r}
source(here("functions/chabaudi_si_opt_lag.R"))
source(here("functions/test2.R"))


# testing with 0.5 parameters of original
orig <- chabaudi_si_opt_lag(parameters_cr = rep(0.5,4),
                                immunity = "tsukushi",
                                parameters = parameters_tsukushi,
                                time_range,
                                df = 3,
                                cue = "I",
                                cue_range = I_range_mid4_log10,
                                solver = "vode",
                                dyn = TRUE,
                                log_cue = "log10") 

# testing lagged model with no lag. Should the same\
test_0 <- chabaudi_si_opt_lag_2(parameters_cr = rep(0.5,4),
                                immunity = "tsukushi",
                                parameters = parameters_tsukushi,
                                time_range,
                                df = 3,
                                cue = "I",
                                cue_range = I_range_mid4_log10,
                                solver = "vode",
                                dyn = TRUE,
                                log_cue = "log10",
                              delay = 0) 

# testing lagged model with 1 day lag
test_1 <- chabaudi_si_opt_lag_2(parameters_cr = rep(0.5,4),
                                immunity = "tsukushi",
                                parameters = parameters_tsukushi,
                                time_range,
                                df = 3,
                                cue = "I",
                                cue_range = I_range_mid4_log10,
                                solver = "vode",
                                dyn = TRUE,
                                log_cue = "log10",
                              delay = 1) 

test_5 <- chabaudi_si_opt_lag(parameters_cr = rep(0.5,4),
                                immunity = "tsukushi",
                                parameters = parameters_tsukushi,
                                time_range,
                                df = 3,
                                cue = "I",
                                cue_range = I_range_mid4_log10,
                                solver = "vode",
                                dyn = TRUE,
                                log_cue = "log10",
                              delay = 5) 

# original and no delay is the same function. Delay leads to same dynamics just later.
ggplot() +
  geom_line(data = orig, aes(x = time, y = value, color = "original")) +
  #geom_line(data = test_0, aes(x = time, y = value, color = "test_0")) +
  #geom_line(data = test_1, aes(x = time, y = value, color = "test_1")) +
  facet_wrap(~variable, scales = "free", ncol = 4) +
  labs(x = "Time post-infection (days)",
         y = "Values",
         color = "") +
  theme_bw() +
  theme(legend.position="bottom")

ggplot() +
  #geom_line(data = orig, aes(x = time, y = value, color = "original")) +
  geom_line(data = test_0, aes(x = time, y = value, color = "test_0")) +
  geom_line(data = test_1, aes(x = time, y = value, color = "test_1")) +
  geom_line(data = test_5, aes(x = time, y = value, color = "test_5")) +
  facet_wrap(~variable, scales = "free", ncol = 4) +
  labs(x = "Time post-infection (days)",
         y = "Values",
         color = "") +
  theme_bw() +
  theme(legend.position="bottom")
```

# testing delayed co=infection model
```{r}
source(here("functions/chabaudi_ci_opt_lag.R"))
source(here("functions/test.R"))

orig <- chabaudi_ci_opt_lag(
  parameters_cr_1 = rep(0.5,4), 
  parameters_cr_2 = rep(0.1,4), 
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = time_range,
  df = 3,
  cue_1 = "I1+I2",
  cue_2 = "I1+I2",
  cue_range_1 = I_range_mid4_log10,
  cue_range_2 = I_range_mid4_log10,
  solver = "vode",
  dyn = TRUE,
  log_cue_1 = "log10",
  log_cue_2 = "log10")

test_0 <- chabaudi_ci_opt_lag_2(
  parameters_cr_1 = rep(0.5,4), 
  parameters_cr_2 = rep(0.1,4), 
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = time_range,
  df = 3,
  cue_1 = "I1+I2",
  cue_2 = "I1+I2",
  cue_range_1 = I_range_mid4_log10,
  cue_range_2 = I_range_mid4_log10,
  solver = "vode",
  dyn = TRUE,
  log_cue_1 = "log10",
  log_cue_2 = "log10",
  delay = 0)

test_1 <- chabaudi_ci_opt_lag_2(
  parameters_cr_1 = rep(0.5,4), 
  parameters_cr_2 = rep(0.1,4), 
  immunity = "tsukushi",
  parameters = parameters_tsukushi,
  time_range = time_range,
  df = 3,
  cue_1 = "I1+I2",
  cue_2 = "I1+I2",
  cue_range_1 = I_range_mid4_log10,
  cue_range_2 = I_range_mid4_log10,
  solver = "vode",
  dyn = TRUE,
  log_cue_1 = "log10",
  log_cue_2 = "log10",
  delay = 1)

# no difference
ggplot() +
  geom_line(data = orig, aes(x = time, y = value, color = strain)) +
  facet_wrap(~variable_alt, scales = "free", ncol = 4) +
  labs(x = "Time post-infection (days)",
         y = "Values",
         color = "") +
  theme_bw() +
  theme(legend.position="bottom")

ggplot() +
  geom_line(data = test_0, aes(x = time, y = value, color = strain)) +
  facet_wrap(~variable_alt, scales = "free", ncol = 4) +
  labs(x = "Time post-infection (days)",
         y = "Values",
         color = "") +
  theme_bw() +
  theme(legend.position="bottom")
```

# rerun co-infection without split dosage
```{r}
cl <- makeCluster(detectCores()); setDefaultCluster(cl = cl)
lag_ci_I_20.opt_tsukushi_I <- co_infection_opt(parameters_cr = rep(0.5, 4),
                 limit = 0.01, 
                 model = chabaudi_ci_opt_lag,
                 immunity = "tsukushi",
                 parameters = parameters_tsukushi,
                 time_range = time_range,
                 df = 3,
                 cue_1 = "I1+I2",
                 cue_2 = "I1+I2",
                 cue_range_1 = I_range_mid4_log10 ,
                 cue_range_2 = I_range_mid4_log10 ,
                 solver = "vode",
                 log_cue_1 = "log10",
                 log_cue_2 = "log10")
stopCluster(cl)

lag_ci_I_20.opt_tsukushi_I
#360.8702 -468.8403 -327.0151 -371.5277
#0.001814631
```


```{r}
# run with 1 day delay (alpha), did not optimize well given that first strain optimization still had lower fitness
cl <- makeCluster(detectCores()); setDefaultCluster(cl = cl)
lag_ci_I_20.opt_tsukushi_I_d1 <- co_infection_opt(parameters_cr = rep(0.5, 4),
                 limit = 0.01, 
                 model = chabaudi_ci_opt_lag,
                 immunity = "tsukushi",
                 parameters = parameters_tsukushi,
                 time_range = time_range,
                 df = 3,
                 cue_1 = "I1+I2",
                 cue_2 = "I1+I2",
                 cue_range_1 = I_range_mid4_log10 ,
                 cue_range_2 = I_range_mid4_log10 ,
                 solver = "vode",
                 log_cue_1 = "log10",
                 log_cue_2 = "log10",
                 delay = 1)
stopCluster(cl)

lag_ci_I_20.opt_tsukushi_I_d1 
```


```{r}
# run with 3 day delay (alpha+alphag)
cl <- makeCluster(detectCores()); setDefaultCluster(cl = cl)
lag_ci_I_20.opt_tsukushi_I_d3 <- co_infection_opt(parameters_cr = rep(0.5, 4),
                 limit = 0.01, 
                 model = chabaudi_ci_opt_lag,
                 immunity = "tsukushi",
                 parameters = parameters_tsukushi,
                 time_range = time_range,
                 df = 3,
                 cue_1 = "I1+I2",
                 cue_2 = "I1+I2",
                 cue_range_1 = I_range_mid4_log10 ,
                 cue_range_2 = I_range_mid4_log10 ,
                 solver = "vode",
                 log_cue_1 = "log10",
                 log_cue_2 = "log10",
                 delay = 3)
stopCluster(cl)

lag_ci_I_20.opt_tsukushi_I_d3[[]]
```

