---
title: "Report10"
output: html_document
---

# load libraries
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(here)
library(npreg)
library(splines)
library(forecast)
library(hts)
library(ggpubr)
library(nlme)
```

# read in experimental data
```{r}
# import in https://academic.oup.com/emph/article/2018/1/127/5045871?login=true
## (2018 published in EMPH)
emph_2018 <- readxl::read_xls(here("experimental_data/Huijben_2018_EMPH.xls"), sheet = 1)

# import in https://onlinelibrary.wiley.com/doi/10.1111/j.1558-5646.2010.01068.x
## (2010 published in Evolution)
evo_2010 <- readxl::read_xls(here("experimental_data/Huijben_2010_evolution.xls"), sheet = 1)

# import in https://onlinelibrary.wiley.com/doi/10.1111/j.1420-9101.2011.02369.x
## (2013 in PLoS pathogen)
plos_2013_1 <- readxl::read_xlsx(here("experimental_data/Huijben_2013_PLoS.xlsx"), sheet = 2)

plos_2013_2 <- readxl::read_xlsx(here("experimental_data/Huijben_2013_PLoS.xlsx"), sheet = 3)

# import in https://onlinelibrary.wiley.com/doi/10.1111/j.1420-9101.2011.02369.x
## (2011 in Journal of Evolutionary Biology)
eseb_2011 <- readxl::read_xls(here("experimental_data/Huijben_2011_eseb.xls"), sheet = 1)

# import in https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3939351/
## (2011 in Journal of American naturalist)
amna_2011 <- readxl::read_xls(here("experimental_data/Pollitt_2011_naturalist.xls"), sheet = 1)

```

# clean data for single infection
```{r}
## for EMPH 2018 study, include only infection series without drugs were R-inoculum is administered by itself, which includes 6, 7, 8, 9, 10. Box 6 has a starting inoculum number of 10^6, which is most similar to other studies. Filtering between day 3-21 because those are the days where we have single day data.
emph_2018_ss.df <- emph_2018 %>% 
  filter(Box %in% seq(6, 10) &
         dplyr::between(Day, 3, 21)) %>% 
  mutate(dose = case_when(
    Box == 6 ~ 10^6,
    Box == 7 ~ 10^5,
    Box == 8 ~ 10^3,
    Box == 9 | Box == 10 ~ 10^1
  )) %>% 
  mutate(strain = "As6p",
         study = "emph2018",
         study_strain = paste0(strain, study),
         id = paste0(study, strain, Box, Mouse, 1),
         RBC = RBC * (10^6)) %>%
  select(day = Day,
         mouse = Mouse, 
         RBC, 
         asex = Rasex,
         gam = Rgam,
         dose,
         strain,
         study,
         study_strain,
         id)

## for 2011 eseb, only day 3-17 data are analyzed because those are the days where gametocyte data are available
eseb_2011_ss.df <- eseb_2011 %>% 
  filter(Clones == "R" & between(Day, 3, 17) &
           Drugs == "N") %>% 
  mutate(dose = 10^6,
         strain = "As8p",
         study = "eseb2011",
         study_strain = paste0(strain, study),
         RBC = RBC*(10^6),
         id = paste0(study, strain, Box, Mouse, 2)) %>% 
  select(day = Day,
         mouse = Mouse,
         RBC,
         asex = R.asex,
         gam = R.gam,
         dose,
         strain,
         study,
         study_strain,
         id)

## for evolution_2010, single infection data for both resistant and susceptible clones are available without drug treatment
evolution_2010_ss.df <- evo_2010 %>% 
  filter(Clone == "R" | Clone == "S") %>%
  filter(between(Day, 3, 21) &
           Drugs == "nodrugs") %>% 
  mutate(asex = R.asex + S.asex,
         gam = R.gam + S.gam,
         dose = 10^6,
         RBC = RBC*(10^6),
         study = "evol2011",
         strain = ifelse(Clone == "R", "As12", "AJ51"),
         study_strain = paste0(strain, "_", study),
         id = paste0(study, strain, Box, Mouse, 3)) %>% 
  select(day = Day,
         mouse = Mouse,
         RBC,
         asex,
         gam,
         dose,
         strain,
         study,
         study_strain,
         id)

## for amnat 2011, get single infection data. Filter out any mice that have missing data. Set negative asexuasl data ot 0
amna_2011_ss.df <-  amna_2011 %>% 
  filter(treat %in% c("AJ", "AS", "ER", "CR", "CW", "DK")) %>% 
  mutate(asex = tot.para - tot.gcyte,
         gam = tot.gcyte,
         dose = 10^6,
         study = "amna_2011",
         RBC = rbc/(10^6),
         study_strain = paste0(treat, "_", study),
         id = paste0(study, treat, div, mouse, 4)) %>% 
  mutate(asex = ifelse(asex < 0, 0, asex)) # sometimes total parasite is less than gametocyte so need to correct for this

### check for NA by groups
amna_na.id <- amna_2011_ss.df %>% 
  filter_at(vars(asex, gam, RBC), all_vars(is.na(.))) %>% 
  distinct(id) %>% 
  select(id)

amna_2011_ss.df2 <- amna_2011_ss.df %>% 
  filter(!(id %in% amna_na.id$id)) %>% 
  select(day,
         mouse,
         RBC,
         asex,
         gam,
         dose,
         strain = treat,
         study,
         study_strain,
         id)

# rbind
exp_ss.df <- rbind(emph_2018_ss.df, eseb_2011_ss.df, evolution_2010_ss.df, amna_2011_ss.df2)

# correct RBC to actual density (in unit of RBC/uL)
exp_ss_dose.df <- exp_ss.df %>% filter(dose == 10^6) 

# add logged parameters
exp_ss_dose.df <- exp_ss_dose.df %>% 
  mutate(log10_asex = log10(asex+1),
         log10_RBC = log10(RBC+1),
         log10_gam = log10(gam+1))

# get dataset with only resistance strain
exp_ss_R.df <- exp_ss_dose.df %>% filter(strain != "AJ51")

```

# data visualization
## filtering by same dosage leads to pretty heterogenous results
```{r}
ggplot(exp_ss.df %>% filter(dose == 10^6)) +
  geom_line(aes(x = day, y = log10(asex), color = id)) +
  facet_wrap(~study) +
  theme_bw()

ggplot(exp_ss.df %>% filter(dose == 10^6)) +
  geom_line(aes(x = day, y = log10(gam), color = id)) +
  facet_wrap(~study) +
  theme_bw()

ggplot(exp_ss.df %>% filter(dose == 10^6)) +
  geom_line(aes(x = day, y = log10(RBC), color = id)) +
  facet_wrap(~study) +
  theme_bw()

## visualizing diffreent dosage
ggplot(exp_ss.df %>% filter(dose != 10^6)) +
  geom_line(aes(x = day, y = asex, color = id)) +
  theme_bw()

ggplot(exp_ss.df %>% filter(dose != 10^6)) +
  geom_line(aes(x = day, y = RBC, color = id)) +
  theme_bw()


## visualization time series variance

```

# Quantifying "noise"
Noise are data in the time series that remain after the data has been detrended (removal of longe term trends, seasonality) and where the data does not exhibit short term autocorrelation. However, what kind of noises are we quantifying here? Is it white noise, where the magnitude of noise is independent from one time point to the next. Or will our noise be more or a ranndom walk, where the extent of noise depends on the extent of noise from a previous time point? I argue that white noise is more appropriate. Certain stochastic processes such as RBC number of gametocyte density do not accumulate. Thus, the extent of stochasticity in one cycle of RBC/gametocyte prdocution do not directly contribute to the extent of stochasticity in the next cycle. You can argue that asexual density could be a random walk process given that the number of merozoite influence the number of merozoite in the next cycle.

How should we group data when we analyze the data? Studies definitely should be segregated. However, the question is whether we analyze by individual mice or with aggregated data of same parasite strain on different mice. The latter option would introduce external noise. Perhaps categorizing the data into an hierarchial structure makes sense, such that we have 3 studies (we could eliminate the non-resistant strain in evo_2011 to cutback on variation), and the individual mice.

# functions for processing data for hts
```{r}
process_hts <- function(df, variable){
  df2 <- df %>%  
  dplyr::group_by(id) %>% 
  dplyr::arrange(day, .by_group = T) %>% 
  dplyr::ungroup() %>% 
  tidyr::spread(key = id, value = variable) %>% 
  janitor::remove_empty(which = "rows") %>% 
  select(c(9:24)) 
  
  df3 <- data.table::data.table(test)[, lapply(.SD, function(x) x[order(is.na(x))])]
  df4 <- zoo::na.trim(test3, is.na = "all")
}
```

# function for calculating SNR and fitting arima model to a bunch of time series data
```{r}
# write function to perform auto arima for groups
bulk_arima <- function(df, variable){
  
  #------------------------#
  # df processing
  #---------------------#
  # get sample size
  n <- n_distinct(df$id)
  
  # split df by unique id
  df.ls <- split(df, df$id)
  
  # get variable data
  data <- lapply(df.ls, function(x) x %>% 
                   arrange(day) %>% 
                   select(variable))

  #--------------------#
  # Detrending using ARIMA
  #--------------------#
  # auto arima
  arima.mod <- lapply(data, auto.arima, stepwise = FALSE, approximation = FALSE, parallel = TRUE)
  
  # store parameters
  coef.ls <- lapply(arima.mod, arimaorder)
  coef.df <- do.call(rbind, coef.ls)
  
  # plot residue plot (should be within)
  residual.ls <- lapply(arima.mod, "residuals")

  # check whether residual is independent using Ljung-Box test (one of the Portmonteau tests. P-value should be above 0.05)
  lb.test <- lapply(lapply(residual.ls, Box.test, type = "Ljung-Box"), function(x) x$p.value)
  ## print message to redo arima if any of the Ljung-Box returns a p-value of below 0.05
  if(any(lb.test < 0.05)){print("Ljung-Box test p-value returns a p-value of below 0.05! Redo arima.")}
  
  #----------------------#
  # Predictability calculation
  #----------------------#
  #----------RMSE (for noise quantification)------------------#
  # calculate RMSE, which is the standard deviation for residue
  rmse.ls <- lapply(residual.ls, sd)
  
  # calculate SNR
  fitted.ls <- lapply(arima.mod, "fitted")
  snr.ls <- mapply(function(x,y) ((abs(x)+1)^2)/((abs(y)+1)^2), fitted.ls, residual.ls)
  snr_mean.ls <- lapply(snr.ls, median)
  
  # turn to df
  res.df <- data.frame(coef = coef.df, 
                       RMSE = unlist(rmse.ls), 
                        SNR = unlist(snr_mean.ls),
                        variable = variable)
  
  return(res.df)

}
```

# bulk AR model
Think like a parasite, while they can integrate lag values in their decision making (correponding to an AR model), they cannot take into past stochasticity into account (correspondint to the MA part). Hence, a more realistic partitioning of time series would be an imperfect forecastor where we are limited to AR model. The residuals will not be stationary (if MA process is involved). However, those are info that parasite cannot extract.
```{r}
# write function to perform auto arima for groups
bulk_ar <- function(df, variable){
  
  #------------------------#
  # df processing
  #---------------------#
  # get sample size
  n <- n_distinct(df$id)
  
  # split df by unique id
  df.ls <- split(df, df$id)
  
  # get variable data
  data <- lapply(df.ls, function(x) x %>% 
                   arrange(day) %>% 
                   select(variable))

  #--------------------#
  # Detrending using ARIMA
  #--------------------#
  # auto arima with AR 
  ar.mod <- lapply(data, auto.arima, d = 0, max.q = 0, stepwise = FALSE, approximation = FALSE, parallel = TRUE)
  
  # store parameters
  coef.ls <- lapply(ar.mod, arimaorder)
  coef.df <- do.call(rbind, coef.ls)
  
  # plot residue plot (should be within)
  residual.ls <- lapply(ar.mod, "residuals")

  #----------------------#
  # Predictability calculation
  #----------------------#
  #----------RMSE (for noise quantification)------------------#
  # calculate RMSE, which is the standard deviation for residue
  rmse.ls <- lapply(residual.ls, sd)
  
  # calculate SNR
  fitted.ls <- lapply(ar.mod, "fitted")
  snr.ls <- mapply(function(x,y) ((abs(x)+1)^2)/((abs(y)+1)^2), fitted.ls, residual.ls)
  snr_mean.ls <- lapply(snr.ls, median, na.rm = T)
  
  # turn to df
  res.df <- data.frame(coef = coef.df, 
                       RMSE = unlist(rmse.ls), 
                        SNR = unlist(snr_mean.ls),
                        variable = variable)
  
  return(res.df)

}
```

# Analyze data
## ARIMA
```{r}
asex <- bulk_arima(exp_ss_dose.df, "asex")
gam <- bulk_arima(exp_ss_dose.df, "gam")
rbc <- bulk_arima(exp_ss_dose.df, "RBC")

log10_asex <- bulk_arima(exp_ss_dose.df, "log10_asex") # not good ARIMA. Pontmorteau sig for 1 sample. However, given we are repeating test for 16 series, this is expected
log10_rbc <- bulk_arima(exp_ss_dose.df, "log10_RBC")
log10_gam <- bulk_arima(exp_ss_dose.df, "log10_gam")


arima.df <- rbind(asex, gam, rbc, log10_asex, log10_rbc, log10_gam)
arima.df <- arima.df %>% 
  mutate(pq = paste0(coef.p, coef.q),
         pdq = paste0(coef.p, coef.d, coef.q))

arima.df
```

# AR
Here restricte to no differencing (malaria parasite cannot detect trends) or moving average processes (cannot detect external non-cues)
```{r}
asex.ar <- bulk_ar(exp_ss_dose.df, "asex")
gam.ar <- bulk_ar(exp_ss_dose.df, "gam")
rbc.ar <- bulk_ar(exp_ss_dose.df, "RBC")

log10_asex.ar <- bulk_ar(exp_ss_dose.df, "log10_asex") 
log10_rbc.ar <- bulk_ar(exp_ss_dose.df, "log10_RBC")
log10_gam.ar <- bulk_ar(exp_ss_dose.df, "log10_gam")

ar.df <- rbind(asex.ar, gam.ar, rbc.ar, log10_asex.ar, log10_rbc.ar, log10_gam.ar)
asex.ar
```


## parameter of ARIMA
Interestingly, RBC tend to be modelled as an AR process with models either having significant correlation of PACF at lag of 1 day or 2 days. Most gametocyte is modelled as an AR process. However, a sizable proportion of gametocyte are not AR process, meaning that gametocyte concentration cannot be predicted from even a day ago. 

Vast majority of gametocytes are best described by its value a day ago with no MA processes. Similar with RBC and asexual iRBC. However, a sizable porportion of RBC can also be predicted by its value 2 days ago (p = 2). Asexual iRBC exhibit some predictability with 2 days lag but not by much.

interestingly, logging stuff actually made tons of processes random walk (0,1,0). It  seems that the message is clear, intercue predictability is highest for RBC, where we can predict future RBC 2 days ahead of time. Inter-cue variability sucks when we log stuff. As we decrease the magnitude of change by logging, the predictive power of 
```{r}
arima.df %>% 
  ggplot() +
  geom_bar(aes(x = as.factor(coef.p), fill = variable), position = position_dodge2(width = 0.9, preserve = "single")) +
  xlab("p") +
  theme_bw()

ar.df %>% 
  ggplot() +
  geom_bar(aes(x = as.factor(coef.p), fill = variable), position = position_dodge2(width = 0.9, preserve = "single")) +
  xlab("p") +
  theme_bw()
``` 

# SNR vs RMSE
RBC tend to exhibit larger RMSE compared to other variables. In general, of course, the higher the concentration a particular variable, the larger the variable. When we look at SNR, however, RBC tend to given us higher SNR. For iRBC, it depends.  

Interestingly, logging variables decreases variability in SNR and RMSE between individual mice. For asexual, logging drastically increases SNR.
```{r}
# plot RMSE vs SNR 
arima.df %>% 
  ggplot() +
  geom_point(aes(x = log10(RMSE), y = log10(SNR), color = variable), size = 2) +
  theme_bw()

fig1b <- ar.df %>% 
  ggplot() +
  geom_point(aes(x = log10(RMSE), y = log10(SNR), color = variable), size = 2) +
  theme_bw()

fig1 <- ggarrange(fig1a, fig1b, common.legend = T, labels = c("A", "B"))
ggsave(here("figures/report10/fig1.png"))

```

# CCF


# checking goodness of fit for repeated measurements
## fit more types of models

```{r}
bulk_gls <- function(variable, df){
  
  # get independent variables
  ind_var <- setdiff(c("asex", "gam", "RBC", "factor(strain)"), variable)
  
  # get list of formula based on independent variable
  form <- unlist(lapply(1:3, function(n) combn(ind_var, n, FUN=function(row) paste0(variable," ~ ", paste0(row, collapse = "+")))))
  
  # run gls modelm (need to nonvectorized because vectorizsed just doesn't want to work with ANOVA)
  mod_1.ls <- list()
  mod_2.ls <- list()
  mod_3.ls <- list()
  for(i in 1:length(form)){
    mod_1.ls[[i]] <- update(gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 1, q = 0)), . ~ ., method = "ML")
    
    mod_2.ls[[i]] <- update(gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 2, q = 0)), . ~ ., method = "ML")
    
    mod_3.ls[[i]] <- update(gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 3, q = 0)), . ~ ., method = "ML")
  }
  
  # rename models
  names(mod_1.ls) <- lapply(form, function(x){paste0(x, "_1")})
  names(mod_2.ls) <- lapply(form, function(x){paste0(x, "_2")})
  names(mod_3.ls) <- lapply(form, function(x){paste0(x, "_3")})
  
  # combine models
  mod.ls <- c(mod_1.ls, mod_2.ls, mod_3.ls)
  
  # run ANOVA
  res <- cbind(eval(parse(text=paste("anova(",paste("mod.ls[[",1:length(mod.ls),"]]",sep="",collapse=","),")"))), name = names(mod.ls))
  
  return(res)

}
```

```{r}
bulk_gls2 <- function(variable, df){
  
  # get independent variables
  ind_var <- setdiff(c("asex", "gam", "RBC"), variable)
  
  # get list of formula based on independent variable
  form <- unlist(lapply(1:2, function(n) combn(ind_var, n, FUN=function(row) paste0(variable," ~ ", paste0(row, collapse = "+")))))
  
  # run gls modelm (need to nonvectorized because vectorizsed just doesn't want to work with ANOVA)
  mod_1.ls <- list()
  mod_2.ls <- list()
  mod_3.ls <- list()
  
  for(i in 1:length(form)){
    mod_1.ls[[i]] <- gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 1, q = 0))
    
    mod_2.ls[[i]] <- gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 2, q = 0))
    
    mod_3.ls[[i]] <- gls(as.formula(form[[i]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = 3, q = 0))
  }
  
  # rename models
  names(mod_1.ls) <- lapply(form, function(x){paste0(x, "_1")})
  names(mod_2.ls) <- lapply(form, function(x){paste0(x, "_2")})
  names(mod_3.ls) <- lapply(form, function(x){paste0(x, "_3")})
  
  # combine models
  mod.ls <- c(mod_1.ls, mod_2.ls, mod_3.ls)
  
  # run ANOVA
  res <- cbind(eval(parse(text=paste("anova(",paste("mod.ls[[",1:length(mod.ls),"]]",sep="",collapse=","),")"))), name = names(mod.ls))
  
  return(res)

}

```

```{r}
bulk_gls3 <- function(variable, df, p){
  # get independent variables
  ind_var <- setdiff(c("asex", "gam", "RBC"), variable)
  
  # get list of p values
  p_seq <- seq(1, p, 1)
  
  # get list of formula based on independent variable
  form <- unlist(list(paste0(variable , "~", ind_var[[1]], "*", ind_var[[2]]),
               paste0(variable , "~", ind_var[[1]], "+", ind_var[[2]]),
               paste0(variable, "~", ind_var[[1]]),
               paste0(variable, "~", ind_var[[2]])))
  
   # run gls modelm (need to nonvectorized because vectorizsed just doesn't want to work with ANOVA)
  mod_1.ls <- list()
  mod_2.ls <- list()
  mod_3.ls <- list()
  mod_4.ls <- list()
  
  # run for loop gls with different p parameters
  mod_1.ls <- lapply(p_seq, function(x){
    gls(as.formula(form[[1]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = x, q = 0))
  })
  
  #--------------------------------#
  mod_2.ls <- lapply(p_seq, function(x){
    gls(as.formula(form[[2]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = x, q = 0))
  })


  #--------------------------------#
  mod_3.ls <- lapply(p_seq, function(x){
    gls(as.formula(form[[3]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = x, q = 0))
  })
  
#--------------------------------#
  mod_4.ls <- lapply(p_seq, function(x){
    gls(as.formula(form[[4]]), data = exp_ss_dose.df, corr = corARMA(form = ~ 1 | id, p = x, q = 0))
  })


  # ANOVA
  anova_1 <- cbind(eval(parse(text=paste("anova(",paste("mod_1.ls[[",1:length(mod_1.ls),"]]",sep="",collapse=","),")"))), name = form[[1]])
  
  anova_2 <- cbind(eval(parse(text=paste("anova(",paste("mod_2.ls[[",1:length(mod_2.ls),"]]",sep="",collapse=","),")"))), name = form[[2]])
  
  anova_3 <- cbind(eval(parse(text=paste("anova(",paste("mod_3.ls[[",1:length(mod_3.ls),"]]",sep="",collapse=","),")"))), name = form[[3]])

  anova_4 <- cbind(eval(parse(text=paste("anova(",paste("mod_4.ls[[",1:length(mod_4.ls),"]]",sep="",collapse=","),")"))), name = form[[4]])

  return(list(anova_1, anova_2, anova_3, anova_4))
  }


```

```{r}
# set p-value of 0.01
bulk_gls3("RBC", exp_ss_dose.df, 3) # consistently p = 3

bulk_gls3("asex", exp_ss_dose.df, 3) # p = 2, 2, 3, 3

bulk_gls3("gam", exp_ss_dose.df, 3)  # p = 1, 3, 1, 3
```

```{r}
rbc_gls <- bulk_gls2("RBC", exp_ss_dose.df)
asex_gls <- bulk_gls2("asex", exp_ss_dose.df)
gam_gls <- bulk_gls2("gam", exp_ss_dose.df)


rbc_gls %>% arrange(AIC) #RBC ~ asex_3
rbc_gls %>% arrange(BIC) #RBC ~ asex_3

asex_gls %>% arrange(AIC) #asex ~ gam_3
asex_gls %>% arrange(BIC) #asex ~ RBC_3

gam_gls %>% arrange(AIC) #gam ~ asex_3
gam_gls %>% arrange(BIC) #gam ~ asex_3
```



