---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(ggplot2)
library(here)
library(deSolve)
library(crone)
library(optimParallel)
library(doParallel)
library(doRNG)
library(arrow)
library(stringr)
library(parallel)
library(ggpubr)
```

# load functions
```{r}
source(here("functions/chabaudi_si_clean.R"))
source(here("functions/chabaudi_ci_clean.R"))

parameters_tsukushi <- c(R1 = 8.89*10^6, # slightly higher
                lambda = 3.7*10^5,
                mu = 0.025, 
                p = 8*10^-6, # doubled form original
                alpha = 1, 
                alphag = 2, 
                beta = 5.721, 
                mum = 48, 
                mug = 4, 
                I0 = 43.85965, 
                Ig0 = 0, 
                a = 150, 
                b = 100, 
                sp = 1,
                psin = 16.69234,
                psiw = 0.8431785,
                phin = 0.03520591, 
                phiw = 550.842,
                iota = 2.18*(10^6),
                rho = 0.2627156)
```

#---------------------------------#
# preparation for si validation
#---------------------------------#
```{r}
# import in si optimum
si_opt.df <- read.csv(here("data/si_opt.csv"))

# import in si cue range
si_cue_range <- read.csv(here("data/cue_range_si.csv"))

# join
si_opt.df2 <- si_opt.df %>% left_join(select(si_cue_range, low, high, by, id), by = "id")

# write csv
write.csv(si_opt.df2, here("job_scripts/si_opt2.csv"))
```


#===================================================#
# Remaking ci reaction norm figure -> overlay single infection
#===================================================#
```{r}
# import in ci dynamics
ci_rn_lim.df <- read.csv(here("data/ci_dyn/ci_rn_lim.csv"))
ci_rug.df <- read.csv(here("data/ci_dyn/ci_rug.csv"))

# import in si dynamics (resimulated with 30 days to ensure resolution)


```

#===================================================#
# what makes a successful competitor? dynamics analysis
#===================================================#
Here, I will simulate the disease curves and time-series cr of co-infection competition that either resulted in a "win" (fitness difference>0) or a lose (fitness difference<0). I will do this for both static competition and optimized competition results




#--------------Static competition-------------------#
```{r}
# import in static competition results
ci_static.df <- read.csv(here("data/ci_static/ci_comp_static.csv"))

# import in optimal coinfection parameters
ci_opt.df <- read.csv(here("data/ci_opt.csv"))
ci_opt.df2 <- ci_opt.df %>% mutate(id = paste0(cue, "_", log))

# join with cue range
cue_range <- read.csv(here("data/cue_range.csv"))

# join cue_1 cue range and cue_2 cue range. also join with optimal parameters
ci_static.df2 <- left_join(ci_static.df, select(cue_range, low_1 = low, high_1 = high, by_1 = by, id),
          by = c("id_1" = "id")) %>% 
  left_join(select(cue_range, low_2 = low, high_2 = high, by_2 = by, id),
            by = c("id_2" ="id")) %>% 
  left_join(select(ci_opt.df2, var1_1 = var1, var2_1 = var2, var3_1 = var3, var4_1 = var4, id), by = c("id_1"= "id")) %>% 
  left_join(select(ci_opt.df2, var1_2 = var1, var2_2 = var2, var3_2 = var3, var4_2 = var4, id), by = c("id_2"= "id")) %>% 
  filter(id_1 != id_2) # filter out same comparison

# filter out wins. 53
ci_static.win <- ci_static.df2 %>% filter(fitness_diff>0)

# filter out loses. 52
ci_static.lose <- ci_static.df2 %>% filter(fitness_diff<0)

```

# function to run ci dynamics
```{r}
ci_dyn <- function(df, win){
  
  # read in df info
  if(stringr::str_detect(df$id_1, "-i")){cue_1 = paste0(df$cue_1, "1")}
  if(stringr::str_detect(df$id_1, "-i", negate = T)){cue_1 = df$cue_1}
  if(stringr::str_detect(df$id_2, "-i")){cue_2 = paste0(df$cue_2, "2")}
  if(stringr::str_detect(df$id_2, "-i", negate = T)){cue_2 = df$cue_2}
  
  # get log
  if(stringr::str_detect(df$id_1, "log")){log_1 = "log10"}
  if(stringr::str_detect(df$id_1, "none")){log_1 = "none"}
  
  if(stringr::str_detect(df$id_2, "log")){log_2 = "log10"}
  if(stringr::str_detect(df$id_2, "none")){log_2 = "none"}
  
  # get parameters
  par_1 <- c(df$var1_1, df$var2_1, df$var3_1, df$var4_1)
  par_2 <- c(df$var1_2, df$var2_2, df$var3_2, df$var4_2)
  
  # get cue range
  cue_range_1 <- seq(df$low_1, df$high_1, by = df$by_1)
  cue_range_2 <- seq(df$low_2, df$high_2, by = df$by_2)
  
  # get unique id
  id <- paste0(df$id_1, df$id_2)
  
  # get dynamics
  dyn <- chabaudi_ci_clean(parameters_cr_1 = par_1,
                  parameters_cr_2 = par_2,
                  immunity = "tsukushi",
                  parameters = parameters_tsukushi,
                  cue_1 = cue_1,
                  cue_2 = cue_2,
                  cue_range_1 = cue_range_1,
                  cue_range_2 = cue_range_2,
                  log_cue_1 = log_1,
                  log_cue_2 = log_2,
                  solver = "vode",
                  time_range = seq(0, 30, 0.001),
                  dyn = T)
  
    dyn2 <- data.frame(dyn, id = rep(id, nrow(dyn)),
                       win = rep(win, nrow(dyn)))
    
    # get filename
    filename <- paste0(id, "_", win, "_dyn.csv")

  write.csv(dyn2, here(paste0("data/ci_static/", filename)))
}
```

# run function
```{r}
ci_static_win.ls <- split(ci_static.win, seq(nrow(ci_static.win)))
ci_static_lose.ls <- split(ci_static.lose, seq(nrow(ci_static.lose)))

mclapply(ci_static_win.ls, function(x){ci_dyn(x, win = "win")}, mc.cores = 6)
mclapply(ci_static_lose.ls, function(x){ci_dyn(x, win = "lose")}, mc.cores = 6)
```

# process shit
```{r}
# convert to parquet you dumbass
convert_parquet <- function(filename){
  temp_csv <- read.csv(filename)
  # get base name without csv
  filename2 <- tools::file_path_sans_ext(basename(filename))
  # write parquet
  write_parquet(temp_csv, sink = here(paste0("data/ci_static/", filename2, ".parquet")))
}

# read in all files
si_static.file <- list.files(path = here("data/ci_static"), pattern = "*_dyn.csv", full.names = T)

# convert to parquet
# mclapply(si_static.file,convert_parquet,  mc.cores = 6)

# import back parquet
si_static_win.file <- list.files(path = here("data/ci_static"), pattern = "*win_dyn.parquet", full.names = T)
si_static_lose.file <- list.files(path = here("data/ci_static"), pattern = "*lose_dyn.parquet", full.names = T)

si_static_win.dyn <- lapply(si_static_win.file, read_parquet)
si_static_lose.dyn <- lapply(si_static_lose.file, read_parquet)

si_static_win.dyn2 <- do.call(rbind, si_static_win.dyn)
write_parquet(si_static_win.dyn2, here("data/ci_static/ci_static_win_final.parquet"))

#si_static_lose.dyn2 <- do.call(rbind, si_static_lose.dyn)
#write_parquet(si_static_lose.dyn2, here("data/ci_static/ci_static_lose_final.parquet"))


# sanity check. look at tau_cum1. good
si_static_win.dyn2 %>% 
  filter(variable == "tau_cum1" | variable == "tau_cum2") %>% 
  group_by(id, variable) %>% 
  summarize(max_fitness = max(value, na.rm = TRUE)) %>% 
  tidyr::spread(variable, max_fitness) %>% 
  mutate(diff = tau_cum1-tau_cum2)

si_static_lose.dyn2 %>% 
  filter(variable == "tau_cum1" | variable == "tau_cum2") %>% 
  group_by(id, variable) %>% 
  summarize(max_fitness = max(value, na.rm = TRUE)) %>% 
  tidyr::spread(variable, max_fitness) %>% 
  mutate(diff = tau_cum1-tau_cum2)

```

# plot time series difference in dynamics for co infection static
```{r}
si_static_win.dyn2 <- read_parquet(here("data/ci_static/ci_static_win_final.parquet"))
# calculate difference between winner and loser
si_static_win.dyn3 <- si_static_win.dyn2 %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value, id_cols = c(time, id)) %>% 
  mutate(total = I1+Ig1)

# process dataframe
si_static_win.dyn4 <- si_static_win.dyn3 %>% 
  mutate(`Gametocyte difference` = G1-G2,
         `Sexual iRBC difference` = Ig1-Ig2,
         `Asexual iRBC difference` = I1-I2,
         `Fitness difference` = tau1-tau2) %>% 
  select(time, id, `Gametocyte difference`, `Sexual iRBC difference`, `Asexual iRBC difference`, `Fitness difference`) %>% 
  tidyr::pivot_longer(!c(time, id)) %>% 
  group_by(id, name) %>% 
  arrange(time, .by_group = T)
  
# plot
ggplot(data = si_static_win.dyn4 %>% filter(row_number() %% 100 ==0), aes(x = time, y = value)) +
  geom_line(aes(group = id), alpha = 0.1) +
  geom_smooth(na.rm = T, span = 0.1, se = 0.95) +
  xlim(0,30)+
  facet_wrap(~name, scales = "free_y") +
  labs(x = "Time (day)", y = "Difference (winner-loser)") +
  geom_hline(yintercept = 0, color = "red") +
  theme_bw()

# explicit plotting of the time-series conversion rate dynamics
ggplot() +
  geom_line(data = si_static_win.dyn3 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1, group = id, color = "Winner"), alpha = 0.05) +
  geom_smooth(data = si_static_win.dyn3 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1, color = "Winner"),na.rm = T, span = 0.1, se = 0.95) +
  geom_line(data = si_static_win.dyn3 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_2, group = id, color = "Loser"), alpha = 0.05) +
  geom_smooth(data = si_static_win.dyn3 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_2, color = "Loser"),na.rm = T, span = 0.1, se = 0.95) +
  xlim(0,30)+
  labs(x = "Time (day)", y = "Conversion rate") +
  scale_color_manual(name= "Competition status",
                     breaks=c("Winner", "Loser"),
                     values=c("Winner" = "blue", "Loser" = "red")) +
  theme_bw()
```

#----------------optimized co-infection parameters---------------------#
```{r}
# import in time series dynamics of optimal co-infection
ci_opt.dyn <- read_parquet(here("data/ci_dyn/ci_opt_40.parquet"))

# import in time series dynamics of non-optimal coinfection
ci_non_opt.dyn <- read_parquet(here("data/ci_dyn/ci_non_opt_40.parquet"))

# function for ci processing
ci_diff_process <- function(df){
  res <- df %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value, id_cols = c(time, label)) %>% 
  mutate(`Gametocyte difference` = G1-G2,
         `Sexual iRBC difference` = Ig1-Ig2,
         `Asexual iRBC difference` = I1-I2,
         `Fitness difference` = tau1-tau2) %>% 
  select(time, label, `Gametocyte difference`, `Sexual iRBC difference`, `Asexual iRBC difference`, `Fitness difference`) %>% 
  tidyr::pivot_longer(!c(time, label)) %>% 
  group_by(label, name) %>% 
  arrange(time, .by_group = T)
  
  return(res)
}

# process df. 
ci_opt.dyn2 <- ci_opt.dyn %>% 
  mutate(infection = "Optimal co-infection") %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value, id_cols = c(time, label, infection)) %>% 
  select(time, label, infection, R, I1, Ig1, G1, tau1, cr_1)
  
ci_non_opt.dyn2 <- ci_non_opt.dyn %>% 
  mutate(infection = "Non-optimal co-infection") %>% 
  tidyr::pivot_wider(names_from = variable, values_from = value, id_cols = c(time, label, infection)) %>% 
  select(time, label, infection, R_n = R, I1_n = I1, Ig1_n = Ig1, G1_n = G1, tau1_n = tau1, cr_1_n = cr_1)

ci_opt.all2 <- left_join(ci_opt.dyn2, ci_non_opt.dyn2, by = c("time", "label")) %>% 
  mutate(`Gametocyte difference` = G1-G1_n,
         `Sexual iRBC difference` = Ig1-Ig1_n,
         `Asexual iRBC difference` = I1-I1_n,
         `Fitness difference` = tau1-tau1_n) %>% 
  select(time, label, `Gametocyte difference`, `Sexual iRBC difference`, `Asexual iRBC difference`, `Fitness difference`) %>% 
  tidyr::pivot_longer(!c(time, label)) %>% 
  group_by(label, name) %>% 
  arrange(time, .by_group = T)
ci_opt.all2

# plot time-series dynamics difference
ggplot(data = ci_opt.all2%>% filter(row_number() %% 100 ==0), aes(x = time, y = value)) +
  geom_line(aes(group = label), alpha = 0.2) +
  geom_smooth(na.rm = T, span = 0.1, se = 0.95) +
  xlim(0,30)+
  facet_wrap(~name, scales = "free_y") +
  labs(x = "Time (day)", y = "Difference (optimal-non-optimal)") +
  geom_hline(yintercept = 0, color = "red") +
  theme_bw()

# plot time-series conversion rate difference
ggplot() +
  geom_line(data = ci_opt.dyn2 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1, group = label, color = "Optimal co-infection"), alpha = 0.15) +
  geom_smooth(data = ci_opt.dyn2%>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1, color = "Optimal co-infection"),na.rm = T, span = 0.1, se = 0.95) +
  geom_line(data = ci_non_opt.dyn2 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1_n, group = label, color = "Non-optimal co-infection"), alpha = 0.15) +
  geom_smooth(data = ci_non_opt.dyn2 %>% 
         filter(row_number() %% 100 ==0), aes(x = time, y = cr_1_n, color = "Non-optimal co-infection"),na.rm = T, span = 0.1, se = 0.95) +
  xlim(0,30)+
  labs(x = "Time (day)", y = "Conversion rate") +
  scale_color_manual(name= "Strategy status",
                     breaks=c("Optimal co-infection", "Non-optimal co-infection"),
                     values=c("Optimal co-infection" = "blue", "Non-optimal co-infection" = "red")) +
  theme_bw()
```

#-----------------optimized co-infection (invasion)----------------------------#
```{r}
# select for best fitness from two study sets: 1 with 0.5x4 starting point and another with original parameter as starting set

# import in the two datasets
ci_invade.file1 <- list.files(path = here("data/ci_invasion/"), pattern = "*.csv", full.names = T)
ci_invade.file2 <- list.files(path = here("data/ci_invasion_2/"),pattern = "*.csv", full.names = T)

ci_invade.df1 <- do.call(rbind, lapply(ci_invade.file1, read.csv))
ci_invade.df2 <- do.call(rbind, lapply(ci_invade.file2, read.csv))

# left_join
ci_invade.df <- left_join(
  select(ci_invade.df1, mut_id, res_id, fitness_1 = fitness),
  select(ci_invade.df2, mut_id, res_id, fitness_2 = fitness),
  by = c("mut_id", "res_id")
  )

# get best fitness
ci_invade.df2 <- ci_invade.df %>% 
  mutate(fitness_final = case_when(
    fitness_1 >= fitness_2 ~ fitness_1,
    fitness_2 > fitness_1 ~ fitness_2
  )) %>% 
  mutate(mut_cue = gsub("\\-.*|\\_.*", "", mut_id),
         mut_log = gsub(".*_", "",  mut_id),
         res_cue = gsub("\\-.*|\\_.*", "", res_id),
         res_log = gsub(".*_", "",  res_id)) %>% 
  mutate(mut_label = ifelse(mut_log == "log", paste(mut_cue, mut_log), mut_cue),
         res_label = ifelse(res_log == "log", paste(res_cue, res_log), res_cue))

write.csv(ci_invade.df2, here("data/ci_invasion_final.csv"))
```

# plot heat map and average
```{r}
ci_invade.pl1 <- ggplot(data = ci_invade.df2, aes(x = res_label, y = mut_label, fill = fitness_final))+
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, space = "Lab", lim = c(-1.8, 1.5), name="Fitness\ndifference") +
  theme_minimal() +  
  theme(legend.position = "left",
  axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  text = element_text(size = 15),
  axis.ticks = element_blank(),
  plot.margin=margin(r = 0)) + 
  labs(x = "Residence cue", y = "Mutant cue") +
  coord_fixed()

# get get average
ci_invade.df3 <- ci_invade.df2 %>% 
  group_by(mut_id) %>% 
  mutate(mean_fitness = mean(fitness_final)) %>% 
  distinct(mut_id, .keep_all = T)

ci_invade.pl2 <- ggplot() +
  geom_bar(data =  ci_invade.df3, aes(y = mut_label, x = mean_fitness), stat = "identity") +
  labs(y = "", x = "Mean fitness\ndifference") +
  theme_bw() +
  theme(plot.margin = margin(l = 0),
       panel.grid.major = element_blank(),
  panel.background = element_blank(),
  text = element_text(size = 15),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank())
  
ci_invade.pl <- ggarrange(ci_invade.pl1, ci_invade.pl2, align = "h", widths = c(1, 0.2))

ggsave(here("figures/report19/ci_invade.png"), width = 10)
```


#====================================================================#
# pairwise comparison of different co-infection competition
#====================================================================#
This graph shows the effects of different types of cue perception (total vs individual, log vs non-logged for both static and semi-dynamic co-infection competition)

#---------------------static competition------------------------------#

```{r}
# import in static competition dynamics
ci_static_win.df <- read_parquet(here("data/ci_static/ci_static_win_final.parquet"))
ci_static_lose.df <- read_parquet(here("data/ci_static/ci_static_lose_final.parquet"))

# combine these two

ci_static_win.df2 <- ci_static_win.df %>% 
  filter(variable == "tau_cum1" | variable == "tau_cum2") %>% 
  filter(time == 30) %>% 
  mutate(cue_1 = stringr::str_split(id, "\\_|log|none") %>% 
           purrr::map_chr(., 1),
         cue_2 = stringr::str_split(id, "\\_|log|none") %>% 
           purrr::map_chr(., 3),
         temp_1 = stringr::str_split(id, "\\_") %>% 
           purrr::map_chr(., 2),
         temp_2 = stringr::str_split(id, "\\_") %>% 
           purrr::map_chr(., 3)) %>% 
  mutate(
    log_1 = case_when(
    str_detect(temp_1, "log") ~ "log10",
    str_detect(temp_1, "none") ~ "none"),
    log_2 = case_when(
      str_detect(temp_2, "log") ~ "log10",
      str_detect(temp_2, "none") ~ "none"),
    )
ci_static_win.df2

write_parquet(ci_static_win.df2, here("data/ci_static/ci_static_win2.parquet"))

ci_static_lose.df2 <- ci_static_lose.df %>% 
  filter(variable == "tau_cum1" | variable == "tau_cum2") %>% 
  filter(time == 30) %>% 
  mutate(cue_1 = stringr::str_split(id, "\\_|log|none") %>% 
           purrr::map_chr(., 1),
         cue_2 = stringr::str_split(id, "\\_|log|none") %>% 
           purrr::map_chr(., 3),
         temp_1 = stringr::str_split(id, "\\_") %>% 
           purrr::map_chr(., 2),
         temp_2 = stringr::str_split(id, "\\_") %>% 
           purrr::map_chr(., 3)) %>% 
  mutate(
    log_1 = case_when(
    str_detect(temp_1, "log") ~ "log10",
    str_detect(temp_1, "none") ~ "none"),
    log_2 = case_when(
      str_detect(temp_2, "log") ~ "log10",
      str_detect(temp_2, "none") ~ "none"),
    )
write_parquet(ci_static_lose.df2, here("data/ci_static/ci_static_lose2.parquet"))

# combine both
ci_static.comp <- rbind(ci_static_win.df2, ci_static_lose.df2)
```

# plot log
```{r}
# get log vs non-log
ci_static_comp.log <- ci_static.comp %>% 
  filter((log_1 == "log10" & log_2 == "none") | (log_1 == "none" & log_2 == "log10")) %>% 
  filter(cue_1 == cue_2) %>% 
  select(strain, value, cue = cue_1, log_1, log_2) %>% 
  mutate(log_alt = case_when(
    strain == 1 & log_1 == "none" ~ "none",
    strain == 2 & log_2 == "log10" ~ "log10"
  ))

log_pair.pl <- ggpaired(ci_static_comp.log, x = "log_alt", y = "value", id = "cue") +
  labs(x = "Cue perception", y = "Fitness")
```

# plot combine vs individual
```{r}
ci_static_comp.comb <- ci_static.comp %>% 
  mutate(cue_alt_1 = case_when(
    stringr::str_detect(cue_1, "-i") ~ gsub("*-i", "", cue_1),
    stringr::str_detect(cue_1, "1+") ~ gsub("*1+.*", "", cue_1)),
    cue_alt_2 = case_when(
    stringr::str_detect(cue_2, "-i") ~ gsub("*-i", "", cue_2),
    stringr::str_detect(cue_2, "1+") ~ gsub("*1+.*", "", cue_2))) %>% 
  filter(cue_alt_1 == cue_alt_2 & log_1 == log_2) %>% 
  mutate(combine = case_when(
    strain == 1 & str_detect(cue_1, "\\+") ~ "Combined",
    strain == 2 & str_detect(cue_2, "-i") ~ "Individual")
  )

comb_pair.pl <- ggpaired(ci_static_comp.comb, x = "combine", y = "value", id = "temp_1") +
  labs(x = "Cue perception", y = "Fitness")
```

```{r}
log_pair.pl
comb_pair.pl
```







